---
title: "Landmark region difference between mean species"
author: "Thomas Guillerme"
bibliography: references.bib
date: "`r Sys.Date()`"
output:
  html_document:
    fig_width: 12
    fig_height: 6
---

```{r, message = FALSE, warning = FALSE}
## Loading the libraries (and installing if necessary)
if(!require(devtools)) install.packages("devtools")
if(!require(knitr)) install.packages("knitr")
if(!require(xtable)) install.packages("xtable")
if(!require(geomorph)) install.packages("geomorph")
if(!require(dispRity)) install.packages("dispRity")
if(!require(landvR)) install_github("TGuillerme/landvR")
source("../Functions/utilities.R")
set.seed(42)
```

Here we test the Procrustes landmarks position variation not between the two extremes (or 95%) of the Procrustes space but between the mean Procrustes shapes between the two genera and the two species in _Lasiorhinus_.

# Preparing the data

First we load the two GPA for either all the three species (to compare _Vombatus vs. Lasiorhinus_) or either for the two _Lasiorhinus_ species (to compare _L. krefftii vs. L. latifrons_).

```{r}
## Reading the full GPA data
path <- "../Data/Processed/"
save_path <- "../Data/Results/"

## The GPA for the cranium and mandible of all the taxa
load(paste0(path, "Wombat", ".Rda"))
all_gpa_cran <- land_data[["cranium"]]
all_gpa_mand <- land_data[["mandible"]]

## Getting the partitions
partitions_cran <- partitions_mand <- list()
for(part in 1:length(unique(all_gpa_cran$landmarkgroups[,2]))) {
    partitions_cran[[part]] <- which(
            all_gpa_cran$landmarkgroups[,2] == unique(all_gpa_cran$landmarkgroups[,2])[part])
    partitions_mand[[part]] <- which(
            all_gpa_mand$landmarkgroups[,2] == unique(all_gpa_mand$landmarkgroups[,2])[part])
}

## The GPA for the cranium and mandible for the Lasiorhinus
load(paste0(path, "wombat_lasiorhinus", ".Rda"))
las_gpa_cran <- land_data[["cranium"]]
las_gpa_mand <- land_data[["mandible"]]

## Creating the groups
## All species
specimens_names <- attr(all_gpa_cran$procrustes$coords, "dimnames")[[3]]
all_groups_cran <- list("Vombatus_ursinus" = grep("CW_", specimens_names),
                "Lasiorhinus" = c(grep("SHNW_", specimens_names), grep("NHNW_", specimens_names)))
specimens_names <- attr(all_gpa_mand$procrustes$coords, "dimnames")[[3]]
all_groups_mand <- list("Vombatus_ursinus" = grep("CW_", specimens_names),
                "Lasiorhinus" = c(grep("SHNW_", specimens_names), grep("NHNW_", specimens_names)))

## Lasiorhinus only
specimens_names <- attr(las_gpa_cran$procrustes$coords, "dimnames")[[3]]
las_groups_cran <- list("Lasiorhinus_latifrons" = grep("SHNW_", specimens_names),
                    "Lasiorhinus_krefftii" = grep("NHNW_", specimens_names))

specimens_names <- attr(las_gpa_mand$procrustes$coords, "dimnames")[[3]]
las_groups_mand <- list("Lasiorhinus_latifrons" = grep("SHNW_", specimens_names),
                    "Lasiorhinus_krefftii" = grep("NHNW_", specimens_names))

```

We can then get the coordinates differences (spherical) between the two pairs of taxa for the cranium and the mandible.


```{r}
## Getting the mean specimens for each group
all_means_cran <- select.procrustes(all_gpa_cran$procrustes, factors = all_groups_cran)
all_means_mand <- select.procrustes(all_gpa_mand$procrustes, factors = all_groups_mand)
las_means_cran <- select.procrustes(las_gpa_cran$procrustes, factors = las_groups_cran)
las_means_mand <- select.procrustes(las_gpa_mand$procrustes, factors = las_groups_mand)

## Get the hypothetical mean based on the PCA and the GPA
get.hypothetical.mean <- function(GPA, PCA, axis) {
    ## Transforming the matrix (generalised from geomorph::plotRefToTarget)
    transform_matrix <- as.vector(t(GPA)) + c(mean(PCA$x[,axis]), rep(0, ncol(PCA$x)-length(axis))) %*% t(PCA$rotation)
    ## Converting into a array
    output <- geomorph::arrayspecs(A = transform_matrix, p = dim(GPA)[1], k = dim(GPA)[2])
    return(output)
}
all_cran_hypothetical <- lapply(all_means_cran, get.hypothetical.mean, PCA = las_gpa_cran$ordination, axis = 1)
all_mand_hypothetical <- lapply(all_means_mand, get.hypothetical.mean, PCA = las_gpa_mand$ordination, axis = 1)
las_cran_hypothetical <- lapply(las_means_cran, get.hypothetical.mean, PCA = las_gpa_cran$ordination, axis = 1)
las_mand_hypothetical <- lapply(las_means_mand, get.hypothetical.mean, PCA = las_gpa_mand$ordination, axis = 1)


## Coordinates differences between pairs of species
differences_list_cran <- list(
    "vomba_lasio" = coordinates.difference(type = "spherical",
                        all_cran_hypothetical$Vombatus_ursinus[,,1], all_cran_hypothetical$Lasiorhinus[,,1]),
    "latif_kreft" = coordinates.difference(type = "spherical",
                        las_cran_hypothetical$Lasiorhinus_latifrons[,,1], las_cran_hypothetical$Lasiorhinus_krefftii[,,1]))
differences_list_mand <- list(
    "vomba_lasio" = coordinates.difference(type = "spherical",
                        all_mand_hypothetical$Vombatus_ursinus[,,1], all_mand_hypothetical$Lasiorhinus[,,1]),
    "latif_kreft" = coordinates.difference(type = "spherical",
                        las_mand_hypothetical$Lasiorhinus_latifrons[,,1], las_mand_hypothetical$Lasiorhinus_krefftii[,,1]))

```


# Running the Procrustes coordinates variation tests

We can then test whether the partitions in each dataset (cranium or mandible) differ in terms of coordinates magnitude change between the two groups (_Vombatus vs. Lasiorhinus_ and _L. krefftii vs. L. latifrons_):

```{r}
## Test wrapper
wrap.rand.test <- function(procrustes_var, partitions, test, rarefaction, verbose = TRUE) {
    if(verbose) message("Running test...")
    if(rarefaction != FALSE) {
        rarefaction <- min(unlist(lapply(partitions, length)))
        test_out <- lapply(partitions, lapply.rand.test, data = procrustes_var[[1]], test = test,
           replicates = 1000, resample = FALSE, rarefaction = rarefaction)
    } else {
        test_out <- lapply(partitions, lapply.rand.test, data = procrustes_var[[1]], test = test,
           replicates = 1000, resample = FALSE)
    }
    if(verbose) message("Done.\n")
    return(test_out)
}

## Running the tests (no rarefaction)
differences_cran <- lapply(differences_list_cran, wrap.rand.test, partitions = partitions_cran,
                           test = area.diff, rarefaction = FALSE)
differences_mand <- lapply(differences_list_mand, wrap.rand.test, partitions = partitions_mand,
                           test = area.diff, rarefaction = FALSE)
overlaps_cran <- lapply(differences_list_cran, wrap.rand.test, partitions = partitions_cran,
                       test = bhatt.coeff, rarefaction = FALSE)
overlaps_mand <- lapply(differences_list_mand, wrap.rand.test, partitions = partitions_mand,
                        test = bhatt.coeff, rarefaction = FALSE)

group_cranium <- list("differences" = differences_cran, "overlaps" = overlaps_cran)
group_mandible <- list("differences" = differences_mand, "overlaps" = overlaps_mand)


## Saving the test results
save(group_cranium, file = paste0(save_path, "Group_cranium_means_hypo.Rda"))
save(group_mandible, file = paste0(save_path, "Group_mandible_means_hypo.Rda"))


## Running the tests (with rarefaction)
differences_cran <- lapply(differences_list_cran, wrap.rand.test, partitions = partitions_cran,
                           test = area.diff, rarefaction = TRUE)
differences_mand <- lapply(differences_list_mand, wrap.rand.test, partitions = partitions_mand,
                           test = area.diff, rarefaction = TRUE)
overlaps_cran <- lapply(differences_list_cran, wrap.rand.test, partitions = partitions_cran,
                       test = bhatt.coeff, rarefaction = TRUE)
overlaps_mand <- lapply(differences_list_mand, wrap.rand.test, partitions = partitions_mand,
                        test = bhatt.coeff, rarefaction = TRUE)

group_cranium <- list("differences" = differences_cran, "overlaps" = overlaps_cran)
group_mandible <- list("differences" = differences_mand, "overlaps" = overlaps_mand)


## Saving the test results
save(group_cranium, file = paste0(save_path, "Group_craniumrarefied_means_hypo.Rda"))
save(group_mandible, file = paste0(save_path, "Group_mandiblerarefied_means_hypo.Rda"))
```
