---
title: "Main analyses"
author: "Vera Weisbecker"
bibliography: references.bib
date: "`r Sys.Date()`"
output:
  html_document:
    fig_width: 12
    fig_height: 6
---

```{r, message = FALSE, warning = FALSE}
## Loading the libraries (and installing if necessary)
if(!require(knitr)) install.packages("knitr")
if(!require(xtable)) install.packages("xtable")
if(!require(geomorph)) install.packages("geomorph")
if(!require(testthat)) install.packages("testthat")
if(!require(dispRity)) install.packages("dispRity")
source("../Functions/utilities.R")
set.seed(42)

## Loading and renaming the land_data objects
load("../Data/Processed/wombat.Rda")
AllWombats <- land_data
load("../Data/Processed/wombat_krefftii.Rda")
NHNW <- land_data
load("../Data/Processed/wombat_latifrons.Rda")
SHNW <- land_data
load("../Data/Processed/wombat_ursinus.Rda")
CW <- land_data

## Read the classifiers
cranium <- read.csv("../Data/Raw/classifier_cranium_wombat.csv")
mandible <- read.csv("../Data/Raw/classifier_mandible_wombat.csv")
## Remove juvenile from mandibles
nhnw_b34 <- which(mandible[,1] == "NHNW_B34")
mandible <- mandible[-nhnw_b34, ]


## Make sure the classifier names match landmark file names
expect_true(all(dimnames(AllWombats$cranium$procrustes$coords)[[3]] == cranium$ID))
expect_true(all(dimnames(AllWombats$mandible$procrustes$coords)[[3]] == mandible$ID))
```

# Preparing the data

## Partition classifiers according to species

<!-- Full_classifier_Table_ is the tables of mandibular or cranial classifiers (e.g. read.csv("../Data/Raw/classifier_cranium_wombat.csv")); spp is a vector of species (e.g. c("CW", "SHNW", "NHNW")) -->

```{r, message = FALSE, warning = FALSE}
## Creating the vectors holding the species names
species <- c("CW", "SHNW", "NHNW")
sp_holders <-c("Common wombat", "SHN wombat", "NHN wombat")

## Assigning the right class to each species (for the cranium)
for (i in 1:length(unique(cranium$Species))) {
    ## Extracting the class for each species   
    class <- cranium[cranium$Species == sp_holders[i],]
    ## Preparing the sp name for giving names to outputs
    sp_name <- paste(species[i], deparse(substitute(cranium)), "class", sep = "_")
    ## Assigning the class to the names
    assign(sp_name, class)
}

## Assigning the right class to each species (for the mandible)
for (i in 1:length(unique(mandible$Species))) {  
    ## Extracting the class for each species   
    class <- mandible[mandible$Species == sp_holders[i],]
    ## Preparing the sp name for giving names to outputs
    sp_name <- paste(species[i], deparse(substitute(mandible)), "class", sep = "_")
    ## Assigning the class to the names
    assign(sp_name, class)
}
```

## Checking for shape outliers

The following session evaluates the shape of the outliers

```{r, message = FALSE, warning = FALSE}
## set up the gridpars for plotRefToTarget to be passed to rgl
points <- gridPar(pt.size = 0.7)
```

This following section detects outliers (the specimen beyond the upper and lower quartiles) and then visually checks whether these individuals have any obvious wrongly placed landmarks.
When multiple individuals are outliers, each are checked in turn (replacing `ID` in `X$Y$procrustes$coords[, , ID]` by the specimens' ID; where `X` and `Y` are respectively the species object and the species dataset).

#### The Common wombat

The cranium:

```{r, message = FALSE, warning = FALSE, eval = FALSE}
## Plot the procrustes distances from mean for specimens to identify the ones beyond
## the upper/lower quartiles
plotOutliers(CW$cranium$procrustes$coords) 
## CW_M11 (individual 14,11 are out)

## Use plotRefToTarget with lollipop graphs to check if outlier has landmark error;
## Replace number according to which individuals have been identified as outliers
plotRefToTarget(CW$cranium$procrustes$consensus, CW$cranium$procrustes$coords[, , 14],
                method = "vector", gridPars = points)
## No evidence of landmarking error
```

The mandible:

```{r, message = FALSE, warning = FALSE, eval = FALSE}
## Plot the procrustes distances from mean for specimens to identify the ones beyond
## the upper/lower quartiles
plotOutliers(CW$mandible$procrustes$coords)
## individuals 20 and 11 are out

## Use plotRefToTarget with lollipop graphs to check if outlier has landmark error;
## Replace number according to which individuals have been identified as outliers
plotRefToTarget(CW$mandible$procrustes$consensus, CW$mandible$procrustes$coords[, , 20],
                method = "vector", gridPars = points)
## No evidence of landmarking error
```

#### The Northern Hairy Nose Wombat

The cranium:

```{r, message = FALSE, warning = FALSE, eval = FALSE}
## Plotting the outliers
plotOutliers(NHNW$cranium$procrustes$coords)
## NHNW_JM8466 (individual 21 is out)

## Plotting the outliers
plotRefToTarget(NHNW$cranium$procrustes$consensus, NHNW$cranium$procrustes$coords[, , 21],
                method = "vector", gridPars=points)
## No evidence of landmarking error
```

The mandible:

```{r, message = FALSE, warning = FALSE, eval = FALSE}
## Plotting the outliers
plotOutliers(NHNW$mandible$procrustes$coords)
## Specimen 11 is out

## Plotting the outliers
plotRefToTarget(NHNW$mandible$procrustes$consensus, NHNW$mandible$procrustes$coords[, , 11],
                method = "vector", gridPars = points)
## No evidence of landmarking error
```

#### The Southern Hairy Nose Wombat

The cranium:

```{r, message = FALSE, warning = FALSE, eval = FALSE}
## Plotting the outliers
plotOutliers(SHNW$cranium$procrustes$coords)
## 21, 20 and 5 are out)

## Plotting the outliers
plotRefToTarget(SHNW$cranium$procrustes$consensus, NHNW$cranium$procrustes$coords[, , 5],
                method = "vector", gridPars = points)
## No evidence of landmarking error
```

The mandible:

```{r, message = FALSE, warning = FALSE, eval = FALSE}
## Plotting the outliers
plotOutliers(SHNW$mandible$procrustes$coords)
## No outliers!
```




# Allometry Analyses

We can run the allometry tests using the `handle.procD.formula` wrapper function:

```{r, message = FALSE, warning = FALSE, result = 'hide'}
## List of all the data
all_data <- list("NHNW" = NHNW, "SHNW" = SHNW, "CW" = CW)

## Placeholder for the results
allom_output <- list()

## Running the allometry tests
for (i in 1:length(all_data)){
    for (k in 1:length(all_data[[1]])){
        ## Result placeholder
        allom_output[i][[k]] <- list()
        
        ## Procrustes allometry test (wrapper around the procD.allometry function)
        allom_output[[i]][[k]] <- handle.procD.formula(coords ~ Csize, all_data[[i]][[k]]$procrustes,
                                                       procD.fun = procD.allometry, iter = 1000,
                                                       print.progress = FALSE)
    }
}

## Naming the results
names(allom_output) <- names(all_data)

## Naming the partitions
for(i in 1:length(all_data)) {
    names(allom_output[[i]]) <- names(all_data[[i]])
}
```

And we can display them as follows:

```{r, echo = FALSE}
kable(allom_output[[1]][[1]]$aov.table, digits = 3,
    caption = "Procrustes ANOVA for size (allometry) for the northern hairy nose wombat's cranium.")
kable(allom_output[[1]][[2]]$aov.table, digits = 3,
   caption = "Procrustes ANOVA for size (allometry) for the northern hairy nose wombat's mandible.")
kable(allom_output[[2]][[1]]$aov.table, digits = 3,
    caption = "Procrustes ANOVA for size (allometry) for the southern hairy nose wombat's cranium.")
kable(allom_output[[2]][[2]]$aov.table, digits = 3,
   caption = "Procrustes ANOVA for size (allometry) for the southern hairy nose wombat's mandible.")
kable(allom_output[[3]][[1]]$aov.table, digits = 3,
              caption = "Procrustes ANOVA for size (allometry) for the common wombat's cranium.")
kable(allom_output[[3]][[2]]$aov.table, digits = 3,
              caption = "Procrustes ANOVA for size (allometry) for the common wombat's mandible.")
```


# 2-Block Partial Least Squares analysis (2BPLS)


## Reduce coordinate datasets to only those that have mandibles and crania

```{r, message = FALSE, warning = FALSE}
## Data list
all_data <- list("NHNW" = NHNW, "SHNW" = SHNW, "CW" = CW)
## Classifiers list
all_classifiers <- list(
                    "NHNW" = list("cranium" = NHNW_cranium_class, "mandible" = NHNW_mandible_class),
                    "SHNW" = list("cranium" = SHNW_cranium_class, "mandible" = SHNW_mandible_class),
                    "CW" = list("cranium" = CW_cranium_class, "mandible" = CW_mandible_class)
                   )

## Checking whether the cranial and mandibular coordinates datasets match
matched_coords <- reduce.check(all_data, all_classifiers)
expect_true(all(unlist(matched_coords[[2]])))
```

## Run 2BPLS

Depending on processor speed, this can take a while, so we recommend just loading the results which are saved in the `Results` folder.
The results can be run using the following code chunk:


```{r, message = FALSE, warning = FALSE, eval = FALSE}
## 2BPLS for the southern hairy nose wombats
SHNWTBPLS <- two.b.pls(matched_coords[[1]]$SHNW$cranium ,matched_coords[[1]]$SHNW$mandible,
                       iter = 1000) 
save(SHNWTBPLS, file = "../Data/Results/SHNWTBLS.Rda")

## 2BPLS for the northern hairy nose wombats
NHNWTBPLS <- two.b.pls(matched_coords[[1]]$NHNW$cranium ,matched_coords[[1]]$NHNW$mandible,
                       iter = 1000) 
save(NHNWTBPLS, file="../Data/Results/NHNWTBLS.Rda")

## 2BPLS for the common wombats
CWTBPLS <- two.b.pls(matched_coords[[1]]$CW$cranium ,matched_coords[[1]]$CW$mandible,
                     iter = 1000) 
save(CWTBPLS, file="../Data/Results/CWTBLS.Rda")
```

Or directly loaded using this one:
 
```{r}
## 2BPLS for the southern hairy nose wombats
load("../Data/Results/SHNWTBLS.Rda")
## 2BPLS for the northern hairy nose wombats
load("../Data/Results/CWTBLS.Rda")
## 2BPLS for the common wombats
load("../Data/Results/NHNWTBLS.Rda")
```

## Subsetting data to run correlations of PLS score with csize and allometry

```{r, message = FALSE, warning = FALSE}
## Getting the list of 2BPLS results
all_TBPLS <- list("NHNW" = NHNWTBPLS, "SHNW" = SHNWTBPLS, "CW" = CWTBPLS)

## Placeholder lists
pls_pca_c_size <- TBPLS_scores <- pc_scores <- centroid_size <- pls_pca <- merged_pls <- list()

## Creating the subsets
for (i in 1:length(all_data)){
 
  ## Placeholder lists
  list() -> 
      pls_pca_c_size [[i]] ->
      TBPLS_scores[[i]] ->
      pc_scores[[i]] -> 
      centroid_size[[i]] ->
      pls_pca [[i]] ->
      merged_pls[[i]]

  ## Create data frames for subsequent merging. 
  for (k in 1:length(all_data$NHNW)){

    ## TBPLS needs to contain i(species), an address for k (x/yscores as per cranium/mandible, and a
    ## designation of how long the first singular vector goes, i.e. the species number); the 5+k i
    ## the "address" of cranial/mandibular scores in the TBPLS data frames
    TBPLS_scores[[i]][[k]] <- as.data.frame(all_TBPLS[[i]][[5+k]]
                                            [c(1:length(dimnames(all_TBPLS[[i]][[5+k]])[[1]]))],
                                            row.names = dimnames(all_TBPLS[[i]][[5+k]])[[1]])
    ## Get the PC scores
    pc_scores[[i]][[k]] <- as.data.frame(all_data[[i]][[k]]$ordination$x[,1])

    ## Get the centroid sizes
    centroid_size[[i]][[k]] <- as.data.frame(all_data[[i]][[k]]$procrustes$Csize)  
    
    ## Combine the pls data
    pls_pca[[i]][[k]] <- merge(TBPLS_scores[[i]][[k]], pc_scores[[i]][[k]],
                               by = "row.names", all.x = TRUE)
    ## Renaming the rows
    row.names(pls_pca[[i]][[k]]) <- pls_pca[[i]][[k]]$Row.names

    ## Combined the merged PLS
    merged_pls[[i]][[k]] <- merge(pls_pca[[i]][[k]], centroid_size[[i]][[k]],
                                  by = "row.names", all.x = TRUE)
    ## Renaming the rows
    row.names(merged_pls[[i]][[k]]) <- merged_pls[[i]][[k]]$Row.names

    ## Remove the two first columns
    merged_pls[[i]][[k]] <- merged_pls[[i]][[k]][,-(1:2)]
    ## Renaming the columns
    colnames(merged_pls[[i]][[k]]) <- c("PLS_Scores", "PC1","Csize")
  }

  ## Add the species names
  names(merged_pls)[[i]] <- names(all_data)[[i]]
  
  ## Add the dataset name (cranium/mandible)
  names(merged_pls[[i]]) <- names(all_data[[1]])
}


## A manual visual check if the above merger is accurate; print the below and the merged_pls
SHNWTBPLS_mandible_Scores <- as.data.frame(SHNWTBPLS$YScores[c(1:19)],
                                           row.names = dimnames(SHNWTBPLS$YScores)[[1]])

SHNWPC1 <- as.data.frame(SHNW$mandible$ordination$x[,1])
SHNWCsize <- as.data.frame(SHNW$mandible$procrustes$Csize)
```

# Run the linear models of PLS scores with PC1 scores and centroid size

```{r, message = FALSE, warning = FALSE}
## Placeholder lists
pca_pls_lm <- c_size_pls_lm <- list()
 
##correlating the pls scores with PC1 scores and centroid sizes for each species and mandible/cranium
for (i in 1:length(merged_pls)){
    ## Placeholders
    pca_pls_lm[[i]] <- c_size_pls_lm[[i]] <- list()
    
    ##this runs the lm analyses for [1] PC scores and [2] Csize
    for (k in 1: length(merged_pls[[1]])){
        pca_pls_lm[[i]][[k]] <- list()
        ##
        pca_pls_lm[[i]][[k]][[1]] <- summary(lm(
                                          merged_pls[[i]][[k]]$PLS_Scores ~ merged_pls[[i]][[k]]$PC1
                                            ))
        ##
        pca_pls_lm[[i]][[k]][[2]] <- summary(lm(
                                        merged_pls[[i]][[k]]$PLS_Scores ~ merged_pls[[i]][[k]]$Csize
                                            ))
       
    }
    ## Renaming the list elements
    names(pca_pls_lm)[[i]] <- names(merged_pls)[[i]]
    names(pca_pls_lm[[i]]) <- names(merged_pls[[1]])
    names(pca_pls_lm[[i]][[k]]) <- c("pls_pca", "PLS_Csize")
}
```


# PCA tests

## PERMANOVA

```{r, message = FALSE, warning = FALSE}
## Getting the ordinated data for all taxa and for the mandible and the cranium
load("../Data/Processed/wombat.Rda")
cranium_PCA <- land_data$cranium$ordination
mandibl_PCA <- land_data$mandible$ordination

## Creating the groups of taxa
groups_cranium <- list("CW"   = grep("CW",   rownames(cranium_PCA$x)),
                       "SHNW" = grep("SHNW", rownames(cranium_PCA$x)),
                       "NHNW" = grep("NHNW", rownames(cranium_PCA$x)))
groups_mandibl <- list("CW"   = grep("CW",   rownames(mandibl_PCA$x)),
                       "SHNW" = grep("SHNW", rownames(mandibl_PCA$x)),
                       "NHNW" = grep("NHNW", rownames(mandibl_PCA$x)))

## Creating the different groups
cranium_pca <- custom.subsets(cranium_PCA$x, groups_cranium)
mandibl_pca <- custom.subsets(mandibl_PCA$x, groups_mandibl)

## Calculating median distance from centroid per group 
## ignore warnings
cranium_anova <- adonis.dispRity(cranium_pca, metric = c(median, centroids))
mandibl_anova <- adonis.dispRity(mandibl_pca, metric = c(median, centroids))
```

The results of the differences per groups:

```{r, eval = TRUE, echo = FALSE, results = 'hide'}
## Combining the two tables
aov_cranium <- as.data.frame(cranium_anova$aov.tab)
rownames(aov_cranium) <- paste(rownames(aov_cranium), "cranium", sep = ".")
aov_mandibl <- as.data.frame(mandibl_anova$aov.tab)
rownames(aov_mandibl) <- paste(rownames(aov_mandibl), "mandible", sep = ".")
table_out <- round(rbind(aov_cranium, aov_mandibl), digits = 3)

caption <- "PERMANOVA results based on the distances of the PCA for both partition (specimens pairwise distances ~ species)"
label <- "PERMANOVA_table"

kable(table_out, digits = 3, caption = caption)

## Save the xtable
textable <- xtable(table_out, caption = caption, label = label)
cat(print(textable, include.rownames = TRUE), file = paste0("../Manuscript/Tables/", label, ".tex"))
```

## Groups sizes

And then we can test the groups pairwise differences (ad hoc)

```{r, message = FALSE, warning = FALSE}
## Calculating median distance from centroid per group 
cranium_med_cent <- dispRity(boot.matrix(cranium_pca), metric = c(median, centroids))
mandibl_med_cent <- dispRity(boot.matrix(mandibl_pca), metric = c(median, centroids))

cranium_test <- test.dispRity(cranium_med_cent, comparisons = "pairwise", test = t.test,
                              correction = "bonferroni")
mandibl_test <- test.dispRity(mandibl_med_cent, comparisons = "pairwise", test = t.test,
                              correction = "bonferroni")
```

The pairwise differences are:

```{r, eval = TRUE, echo = FALSE, results = 'hide'}
## Combining the two tables
cranium_tests <- do.call(cbind, cranium_test)
rownames(cranium_tests) <- paste(gsub(" : ", ":", rownames(cranium_tests)), "cranium", sep = "-")
mandibl_tests <- do.call(cbind, mandibl_test)
rownames(mandibl_tests) <- paste(gsub(" : ", ":", rownames(mandibl_tests)), "mandible", sep = "-")

table_out <- round(rbind(cranium_tests, mandibl_tests), digits = 3)

caption <- "Pairwise t-tests results based on the median centroid distance from each group (with Bonferroni correction)."
label <- "PCA_t_tests_centroid"

kable(table_out, digits = 3, caption = caption)

## Save the xtable
textable <- xtable(table_out, caption = caption, label = label)
cat(print(textable, include.rownames = TRUE), file = paste0("../Manuscript/Tables/", label, ".tex"))
```


```{r, eval = TRUE, echo = FALSE, fig.height = 6, fig.width = 12}
## Plotting the results
par(mfrow = c(1, 2))
plot(cranium_med_cent, main = "Median distances from centroid per groups (cranium)", xlab = "Size (distance from centroid)")
plot(mandibl_med_cent, main = "Median distances from centroid per groups (mandible)", xlab = "", ylab = "")

## saving the results
pdf(file = paste0("../Manuscript/Figures/", label, ".pdf"), width = 12, height = 6)
par(mfrow = c(1, 2))
plot(cranium_med_cent, main = "Median distances from centroid per groups (cranium)", xlab = "Size (distance from centroid)")
plot(mandibl_med_cent, main = "Median distances from centroid per groups (mandible)", xlab = "", ylab = "")
```


## Distances from the centre (per groups)

And then we can test the groups pairwise differences (ad hoc)

```{r, message = FALSE, warning = FALSE}
## Calculating median distance from centroid per group 
cranium_dis_cent <- dispRity(boot.matrix(cranium_pca), metric = c(median, centroids), centroid = 0)
mandibl_dis_cent <- dispRity(boot.matrix(mandibl_pca), metric = c(median, centroids), centroid = 0)

cranium_test <- test.dispRity(cranium_dis_cent, comparisons = "pairwise", test = t.test,
                              correction = "bonferroni")
mandibl_test <- test.dispRity(mandibl_dis_cent, comparisons = "pairwise", test = t.test,
                              correction = "bonferroni")
```

The pairwise differences are:

```{r, eval = TRUE, echo = FALSE, results = 'hide'}
## Combining the two tables
cranium_tests <- do.call(cbind, cranium_test)
rownames(cranium_tests) <- paste(gsub(" : ", ":", rownames(cranium_tests)), "cranium", sep = "-")
mandibl_tests <- do.call(cbind, mandibl_test)
rownames(mandibl_tests) <- paste(gsub(" : ", ":", rownames(mandibl_tests)), "mandible", sep = "-")

table_out <- round(rbind(cranium_tests, mandibl_tests), digits = 3)

caption <- "Pairwise t-tests results based on the median distances the centre for each group (with Bonferroni correction)."
label <- "PCA_t_tests_centre"

kable(table_out, digits = 3, caption = caption)

## Save the xtable
textable <- xtable(table_out, caption = caption, label = label)
cat(print(textable, include.rownames = TRUE), file = paste0("../Manuscript/Tables/", label, ".tex"))
```


```{r, eval = TRUE, echo = FALSE, fig.height = 6, fig.width = 12}
## Plotting the results
par(mfrow = c(1, 2))
plot(cranium_dis_cent, main = "Median distances from the centre per groups (cranium)", xlab = "Position (distance from centre)")
plot(mandibl_dis_cent, main = "Median distances from the centre per groups (mandible)", xlab = "", ylab = "")

## saving the results
pdf(file = paste0("../Manuscript/Figures/", label, ".pdf"), width = 12, height = 6)
par(mfrow = c(1, 2))
plot(cranium_dis_cent, main = "Median distances from the centre per groups (cranium)", xlab = "Position (distance from centre)")
plot(mandibl_dis_cent, main = "Median distances from the centre per groups (mandible)", xlab = "", ylab = "")
```
