---
title: "Standard analyses"
author: "Vera Weisbecker"
bibliography: references.bib
date: "`r Sys.Date()`"
output:
  html_document:
    fig_width: 12
    fig_height: 6
---

```{r, message = FALSE, warning = FALSE}
## Loading the libraries (and installing if necessary)
if(!require(knitr)) install.packages("knitr"); library(knitr)
if(!require(xtable)) install.packages("xtable"); library(xtable)
if(!require(geomorph)) install.packages("geomorph"); library(geomorph)
if(!require(dispRity)) install.packages("dispRity");library(dispRity)
source("../Functions/utilities.R")
set.seed(42)

## Loading and renaming the land_data objects
load("../Data/Processed/wombat.Rda")
AllWombats <- land_data
load("../Data/Processed/wombat_krefftii.Rda")
NHNW <- land_data
load("../Data/Processed/wombat_latifrons.Rda")
SHNW <- land_data
load("../Data/Processed/wombat_ursinus.Rda")
CW <- land_data

## Read the classifiers
cranium <- read.csv("../Data/Raw/classifier_cranium_wombat.csv")
mandible <- read.csv("../Data/Raw/classifier_mandible_wombat.csv")
## Remove juvenile from mandibles
nhnw_b34 <- which(mandible[,1] == "NHNW_B34")
mandible <- mandible[-nhnw_b34, ]
```

# Preparing the data


## Checking for shape outliers

The following session evaluates the shape of the outliers

```{r, message = FALSE, warning = FALSE}
## set up the gridpars for plotRefToTarget to be passed to rgl
points <- gridPar(pt.size = 0.7)
```

This following section detects outliers (the specimen beyond the upper and lower quartiles) and then visually checks whether these individuals have any obvious wrongly placed landmarks.
When multiple individuals are outliers, each are checked in turn (replacing `ID` in `X$Y$procrustes$coords[, , ID]` by the specimens' ID; where `X` and `Y` are respectively the species object and the species dataset).

#### The Common wombat

The cranium:

```{r, message = FALSE, warning = FALSE, eval = FALSE}
## Plot the procrustes distances from mean for specimens to identify the ones beyond
## the upper/lower quartiles
plotOutliers(CW$cranium$procrustes$coords) 
## CW_M11 (individual 14,11 are out)

## Use plotRefToTarget with lollipop graphs to check if outlier has landmark error;
## Replace number according to which individuals have been identified as outliers
plotRefToTarget(CW$cranium$procrustes$consensus, CW$cranium$procrustes$coords[, , 14],
                method = "vector", gridPars = points)
## No evidence of landmarking error
```

The mandible:

```{r, message = FALSE, warning = FALSE, eval = FALSE}
## Plot the procrustes distances from mean for specimens to identify the ones beyond
## the upper/lower quartiles
plotOutliers(CW$mandible$procrustes$coords)
## individuals 20 and 11 are out

## Use plotRefToTarget with lollipop graphs to check if outlier has landmark error;
## Replace number according to which individuals have been identified as outliers
plotRefToTarget(CW$mandible$procrustes$consensus, CW$mandible$procrustes$coords[, , 20],
                method = "vector", gridPars = points)
## No evidence of landmarking error
```

#### The Northern Hairy Nose Wombat

The cranium:

```{r, message = FALSE, warning = FALSE, eval = FALSE}
## Plotting the outliers
plotOutliers(NHNW$cranium$procrustes$coords)
## NHNW_JM8466 (individual 21 is out)

## Plotting the outliers
plotRefToTarget(NHNW$cranium$procrustes$consensus, NHNW$cranium$procrustes$coords[, , 21],
                method = "vector", gridPars=points)
## No evidence of landmarking error
```

The mandible:

```{r, message = FALSE, warning = FALSE, eval = FALSE}
## Plotting the outliers
plotOutliers(NHNW$mandible$procrustes$coords)
## Specimen 11 is out

## Plotting the outliers
plotRefToTarget(NHNW$mandible$procrustes$consensus, NHNW$mandible$procrustes$coords[, , 11],
                method = "vector", gridPars = points)
## No evidence of landmarking error
```

#### The Southern Hairy Nose Wombat

The cranium:

```{r, message = FALSE, warning = FALSE, eval = FALSE}
## Plotting the outliers
plotOutliers(SHNW$cranium$procrustes$coords)
## 21, 20 and 5 are out)

## Plotting the outliers
plotRefToTarget(SHNW$cranium$procrustes$consensus, NHNW$cranium$procrustes$coords[, , 5],
                method = "vector", gridPars = points)
## No evidence of landmarking error
```

The mandible:

```{r, message = FALSE, warning = FALSE, eval = FALSE}
## Plotting the outliers
plotOutliers(SHNW$mandible$procrustes$coords)
## No outliers!
```

# disparity comparison betweeen wombat species

```{r, message = FALSE, warning = FALSE}
## Create gdfs with species for the data frames
cranium_gdf  <- geomorph.data.frame(coords = AllWombats$cranium$procrustes$coords,
                                    Csize = AllWombats$cranium$procrustes$Csize,
                                    species = as.factor(cranium$Species))
mandible_gdf <- geomorph.data.frame(coords = AllWombats$mandible$procrustes$coords,
                                    Csize = AllWombats$mandible$procrustes$Csize,
                                    species = as.factor(mandible$Species))

## Running the analyses on the gdf objects (Disparity within groups compared to that group's mean)
cranium_disparity <- morphol.disparity(coords ~ species, groups = ~species, iter = 1000, seed = 42,
                                       data = cranium_gdf, print.progress = FALSE)

## Disparity within groups compared to that group's mean
mandible_disparity <- morphol.disparity(coords ~ species, groups = ~species, iter = 1000, seed = 42,
                                        data = mandible_gdf, print.progress = FALSE)
```

And we can display the results as follows:

```{r, echo = FALSE}
## Cranium disparity values
kable(data.frame("Procrustes variance" = cranium_disparity[[1]]), digits = 5, caption = "Cranium disparity (procrustes variance) per groups.")
## Cranium disparity differences
kable(cranium_disparity[[2]], digits = 5, caption = "Cranium pairwise absolute differences between variances between groups.")
## Cranium disparity differences p
kable(cranium_disparity[[3]], digits = 5, caption = "Cranium pairwise absolute differences between variances between groups (p-values).")

## Mandible disparity values
kable(data.frame("Procrustes variance" = mandible_disparity[[1]]), digits = 5, caption = "Mandible disparity (procrustes variance) per groups.")
## Mandible disparity differences
kable(mandible_disparity[[2]], digits = 5, caption = "Mandible pairwise absolute differences between variances between groups.")
## Mandible disparity differences p
kable(mandible_disparity[[3]], digits = 5, caption = "Mandible pairwise absolute differences between variances between groups (p-values).")
```

#Are wombats sexually dimorphic?


## Make the sex subsets

The following analysis don't contain the northern hairy nose wombats because there is no variation in location and no sex assignment in the dataset.

```{r, message = FALSE, warning = FALSE}
## Making the classifiers list
all_data <- list("SHNW" = SHNW, "CW" = CW)
all_classifiers <- list("SHNW" = list("cranium" = SHNW_cranium_class,
                                      "mandible" = SHNW_mandible_class),
                          "CW" = list("cranium" = CW_cranium_class,
                                      "mandible" = CW_mandible_class)
                        )
```

Creating the sex subsets and the centroid size per sex subsets:

```{r}
## Results placeholder
coords_sex_output <- c_size_sex_output <- list()

## Getting the coordinates for specimens with Geolocation
for (i in 1:length(all_data)){
    ## Result placeholder
    list() ->
        coords_sex_output[[i]] ->
        c_size_sex_output[[i]]

    ## Getting the coordinates for each classifier
    for (k in 1:length(all_classifiers[[i]])){
        
        ## Sex coordinates
        coords_sex_output[[i]][[k]] <- all_data[[i]][[k]]$procrustes$coords[ , ,
                                       !is.na(all_classifiers[[i]][[k]]$Sex)]
        ## Sex centroid size
        c_size_sex_output[[i]][[k]] <- all_data[[i]][[k]]$procrustes$Csize[
                                       !is.na(all_classifiers[[i]][[k]]$Sex)]

    }
    ## Naming the element
    names(all_data[[i]]) ->
      names(coords_sex_output[[i]]) ->
      names(c_size_sex_output[[i]])
}

names(all_data) ->
  names(coords_sex_output) ->
  names(c_size_sex_output)  
```

Creating the coordinates subsets and the centroid size subsets for geolocation and sex:

```{r}
## Results placeholder

sex_gdfs <- list()

## Creating the geomorph dataframe for both species and datasets
for (i in 1:length(coords_geo_output)){
    ## Results placeholder
    
    sex_gdfs[[i]] <- list()
    
    for (k in 1:length(all_classifiers[[i]])){
      
      ## Sex data frame object
      sex_gdfs[[i]][[k]] <- geomorph.data.frame(
                          coords = coords_sex_output[[i]][[k]],
                          Csize = c_size_sex_output[[i]][[k]],
                          Sex= all_classifiers[[i]][[k]]$Sex[!is.na(all_classifiers[[i]][[k]]$Sex)]
                          )
  }
  ## Naming the elements
  names(sex_gdfs[[i]]) <- names(all_data[[i]])
}
 names(sex_gdfs) <- names(all_data) 

## Making sure the dimension names are correct;requires testthat
#expect_equal(match(dimnames(geo_gdfs$SHNW$cranium$coords)[[3]],names(geo_gdfs$SHNW$cranium$Csize)),
#             seq(1:length(names(geo_gdfs$SHNW$cranium$Csize))))
```


## Running procD.lm analyses

We can run the three following linear models for each species and each datasets:

 * coordinates ~ sex + centroid size
 * centroid size ~ sex

```{r, message = FALSE, warning = FALSE}
## Results placeholder
coord_sex_lm <- csize_sex_lm <- list()

## Running all the linear models
for (i in 1:length(geo_gdfs)){
    ## Results placeholder
   coord_sex_lm[[i]] <- csize_sex_lm[[i]] <- list()
        
    for (k in 1:length(geo_gdfs[[2]])){
        
        ## Running the coordinates ~ sex + centroid size models
        coord_sex_lm[[i]][[k]] <- procD.lm(coords ~ Sex , print.progress = FALSE,
                                           data = sex_gdfs[[i]][[k]], iter = 1000)
        ## Running the centroid size ~ sex models
        csize_sex_lm[[i]][[k]] <- procD.lm(Csize ~ Sex, print.progress = FALSE,
                                           data = sex_gdfs[[i]][[k]], iter = 1000)
    }
    ## Renaming the models
    
        names(coord_sex_lm[[i]]) ->
        names(csize_sex_lm[[i]])
  }

 names(coord_sex_lm) <- names(csize_sex_lm) 
```


And we can display them as follows:

```{r, echo = FALSE}

## coordinates ~ sex + centroid size
kable(coord_sex_lm[[1]][[1]]$aov.table, digits = 3,
    caption = "Procrustes linear model for coordinates as a function of sex and centroid size for the southern hairy nose wombat's cranium.")
kable(coord_sex_lm[[1]][[2]]$aov.table, digits = 3,
    caption = "Procrustes linear model for coordinates as a function of sex and centroid size for the southern hairy nose wombat's mandible.")
kable(coord_sex_lm[[2]][[1]]$aov.table, digits = 3,
    caption = "Procrustes linear model for coordinates as a function of sex and centroid size for the common wombat's cranium.")
kable(coord_sex_lm[[2]][[2]]$aov.table, digits = 3,
    caption = "Procrustes linear model for coordinates as a function of sex and centroid size for the common wombat's mandible.")

## centroid size ~ sex 
kable(csize_sex_lm[[1]][[1]]$aov.table, digits = 3,
    caption = "Procrustes linear model for centroid size as a function of sex for the southern hairy nose wombat's cranium.")
kable(csize_sex_lm[[1]][[2]]$aov.table, digits = 3,
    caption = "Procrustes linear model for centroid size as a function of sex for the southern hairy nose wombat's mandible.")
kable(csize_sex_lm[[2]][[1]]$aov.table, digits = 3,
    caption = "Procrustes linear model for centroid size as a function of sex for the common wombat's cranium.")
kable(csize_sex_lm[[2]][[2]]$aov.table, digits = 3,
    caption = "Procrustes linear model for centroid size as a function of sex for the common wombat's mandible.")
```


# Allometry Analyses

We can run the allometry tests using the `handle.procD.formula` wrapper function:

```{r, message = FALSE, warning = FALSE, result = 'hide'}
## List of all the data
all_data <- list("NHNW" = NHNW, "SHNW" = SHNW, "CW" = CW)

## Placeholder for the results
allom_output <- list()

## Running the allometry tests
for (i in 1:length(all_data)){
    for (k in 1:length(all_data[[1]])){
        ## Result placeholder
        allom_output[i][[k]] <- list()
        
        ## Procrustes allometry test (wrapper around the procD.allometry function)
        allom_output[[i]][[k]] <- handle.procD.formula(coords ~ Csize, all_data[[i]][[k]]$procrustes,
                                                       procD.fun = procD.lm, iter = 1000,
                                                       print.progress = FALSE)
    }
}

## Naming the results
names(allom_output) <- names(all_data)

## Naming the partitions
for(i in 1:length(all_data)) {
    names(allom_output[[i]]) <- names(all_data[[i]])
}
```

And we can display them as follows:

```{r, echo = FALSE}
kable(allom_output[[1]][[1]]$aov.table, digits = 3,
    caption = "Procrustes ANOVA for size (allometry) for the northern hairy nose wombat's cranium.")
kable(allom_output[[1]][[2]]$aov.table, digits = 3,
   caption = "Procrustes ANOVA for size (allometry) for the northern hairy nose wombat's mandible.")
kable(allom_output[[2]][[1]]$aov.table, digits = 3,
    caption = "Procrustes ANOVA for size (allometry) for the southern hairy nose wombat's cranium.")
kable(allom_output[[2]][[2]]$aov.table, digits = 3,
   caption = "Procrustes ANOVA for size (allometry) for the southern hairy nose wombat's mandible.")
kable(allom_output[[3]][[1]]$aov.table, digits = 3,
              caption = "Procrustes ANOVA for size (allometry) for the common wombat's cranium.")
kable(allom_output[[3]][[2]]$aov.table, digits = 3,
              caption = "Procrustes ANOVA for size (allometry) for the common wombat's mandible.")
```

#Allometry between species

```{r}
#Cranium

##turn into gdf frame 
gdf=geomorph.data.frame(coords=AllWombats$cranium$procrustes$coords, Csize=AllWombats$cranium$procrustes$Csize)

##run analysis
AllomCranium=procD.allometry(coords~Csize, data=gdf, method="RegScore")

#Mandible

##turn into gdf frame 
gdf=geomorph.data.frame(coords=AllWombats$mandible$procrustes$coords, Csize=AllWombats$mandible$procrustes$Csize)

##run analysis
AllomMandible=procD.allometry(coords~Csize, data=gdf, method="RegScore")
```

## Allometry analysis without the patches

We can also run the allometry analysis on the data without the patches


```{r, message = FALSE, warning = FALSE, result = 'hide'}
## Placeholder for the results
allom_output_no_patch <- list()

## Running the allometry tests
for (i in 1:length(all_data)){
    for (k in 1:length(all_data[[1]])){
        ## Result placeholder
        allom_output_no_patch[i][[k]] <- list()
        
        ## Procrustes allometry test (wrapper around the procD.allometry function)
        allom_output_no_patch[[i]][[k]] <- handle.procD.formula(coords ~ Csize,
                                                       all_data[[i]][[k]]$procrustes_no_patches,
                                                       procD.fun = procD.lm, iter = 1000,
                                                       print.progress = FALSE)
    }
}

## Naming the results
names(allom_output_no_patch) <- names(all_data)

## Naming the partitions
for(i in 1:length(all_data)) {
    names(allom_output_no_patch[[i]]) <- names(all_data[[i]])
}
```

And we can display them as follows:

```{r, echo = FALSE}
kable(allom_output_no_patch[[1]][[1]]$aov.table, digits = 3,
    caption = "Procrustes ANOVA for size (allometry) for the northern hairy nose wombat's cranium (no patches).")
kable(allom_output_no_patch[[1]][[2]]$aov.table, digits = 3,
   caption = "Procrustes ANOVA for size (allometry) for the northern hairy nose wombat's mandible (no patches).")
kable(allom_output_no_patch[[2]][[1]]$aov.table, digits = 3,
    caption = "Procrustes ANOVA for size (allometry) for the southern hairy nose wombat's cranium (no patches).")
kable(allom_output_no_patch[[2]][[2]]$aov.table, digits = 3,
   caption = "Procrustes ANOVA for size (allometry) for the southern hairy nose wombat's mandible (no patches).")
kable(allom_output_no_patch[[3]][[1]]$aov.table, digits = 3,
              caption = "Procrustes ANOVA for size (allometry) for the common wombat's cranium (no patches).")
kable(allom_output_no_patch[[3]][[2]]$aov.table, digits = 3,
              caption = "Procrustes ANOVA for size (allometry) for the common wombat's mandible (no patches).")
```

#Allometry between species (without the patches)

```{r}
#Cranium

##turn into gdf frame 
gdf=geomorph.data.frame(coords=AllWombats$cranium$procrustes_no_patches$coords,
                        Csize=AllWombats$cranium$procrustes_no_patches$Csize)

##run analysis
AllomCranium=procD.lm(coords~Csize, data=gdf, method="RegScore")

#Mandible

##turn into gdf frame 
gdf=geomorph.data.frame(coords=AllWombats$mandible$procrustes_no_patches$coords,
                        Csize=AllWombats$mandible$procrustes_no_patches$Csize)

##run analysis
AllomMandible=procD.lm(coords~Csize, data=gdf, method="RegScore")
```


# 2-Block Partial Least Squares analysis (2BPLS)

## Partition classifiers according to species

<!-- Full_classifier_Table_ is the tables of mandibular or cranial classifiers (e.g. read.csv("../Data/Raw/classifier_cranium_wombat.csv")); spp is a vector of species (e.g. c("CW", "SHNW", "NHNW")) -->

```{r, message = FALSE, warning = FALSE}
## Creating the vectors holding the species names
species <- c("CW", "SHNW", "NHNW")
sp_holders <-c("Common wombat", "SHN wombat", "NHN wombat")

## Assigning the right class to each species (for the cranium)
for (i in 1:length(unique(cranium$Species))) {
    ## Extracting the class for each species   
    class <- cranium[cranium$Species == sp_holders[i],]
    ## Preparing the sp name for giving names to outputs
    sp_name <- paste(species[i], deparse(substitute(cranium)), "class", sep = "_")
    ## Assigning the class to the names
    assign(sp_name, class)
}

## Assigning the right class to each species (for the mandible)
for (i in 1:length(unique(mandible$Species))) {  
    ## Extracting the class for each species   
    class <- mandible[mandible$Species == sp_holders[i],]
    ## Preparing the sp name for giving names to outputs
    sp_name <- paste(species[i], deparse(substitute(mandible)), "class", sep = "_")
    ## Assigning the class to the names
    assign(sp_name, class)
}
```

## Reduce coordinate datasets to only those that have mandibles and crania

```{r, message = FALSE, warning = FALSE}
## Data list
all_data <- list("NHNW" = NHNW, "SHNW" = SHNW, "CW" = CW)
## Classifiers list
all_classifiers <- list(
                    "NHNW" = list("cranium" = NHNW_cranium_class, "mandible" = NHNW_mandible_class),
                    "SHNW" = list("cranium" = SHNW_cranium_class, "mandible" = SHNW_mandible_class),
                    "CW" = list("cranium" = CW_cranium_class, "mandible" = CW_mandible_class)
                   )

## Checking whether the cranial and mandibular coordinates datasets match
matched_coords <- reduce.check(all_data, all_classifiers)
```

## Run 2BPLS

Depending on processor speed, this can take a while, so we recommend just loading the results which are saved in the `Results` folder.
The results can be run using the following code chunk:


```{r, message = FALSE, warning = FALSE, eval = FALSE}
## 2BPLS for the southern hairy nose wombats
SHNWTBPLS <- two.b.pls(matched_coords[[1]]$SHNW$cranium ,matched_coords[[1]]$SHNW$mandible,
                       iter = 1000) 
save(SHNWTBPLS, file = "../Data/Results/SHNWTBLS.Rda")

## 2BPLS for the northern hairy nose wombats
NHNWTBPLS <- two.b.pls(matched_coords[[1]]$NHNW$cranium ,matched_coords[[1]]$NHNW$mandible,
                       iter = 1000) 
save(NHNWTBPLS, file="../Data/Results/NHNWTBLS.Rda")

## 2BPLS for the common wombats
CWTBPLS <- two.b.pls(matched_coords[[1]]$CW$cranium ,matched_coords[[1]]$CW$mandible,
                     iter = 1000) 
save(CWTBPLS, file="../Data/Results/CWTBLS.Rda")
```

Or directly loaded using this one:
 
```{r}
## 2BPLS for the southern hairy nose wombats
load("../Data/Results/SHNWTBLS.Rda")
## 2BPLS for the northern hairy nose wombats
load("../Data/Results/CWTBLS.Rda")
## 2BPLS for the common wombats
load("../Data/Results/NHNWTBLS.Rda")
```

## Subsetting data to run correlations of PLS score with csize and allometry

```{r, message = FALSE, warning = FALSE}
## Getting the list of 2BPLS results
all_TBPLS <- list("NHNW" = NHNWTBPLS, "SHNW" = SHNWTBPLS, "CW" = CWTBPLS)

## Placeholder lists
pls_pca_c_size <- TBPLS_scores <- pc_scores <- centroid_size <- pls_pca <- merged_pls <- list()

## Creating the subsets
for (i in 1:length(all_data)){
 
  ## Placeholder lists
  list() -> 
      pls_pca_c_size [[i]] ->
      TBPLS_scores[[i]] ->
      pc_scores[[i]] -> 
      centroid_size[[i]] ->
      pls_pca [[i]] ->
      merged_pls[[i]]

  ## Create data frames for subsequent merging. 
  for (k in 1:length(all_data$NHNW)){

    ## TBPLS needs to contain i(species), an address for k (x/yscores as per cranium/mandible, and a
    ## designation of how long the first singular vector goes, i.e. the species number); the 5+k i
    ## the "address" of cranial/mandibular scores in the TBPLS data frames
    TBPLS_scores[[i]][[k]] <- as.data.frame(all_TBPLS[[i]][[5+k]]
                                            [c(1:length(dimnames(all_TBPLS[[i]][[5+k]])[[1]]))],
                                            row.names = dimnames(all_TBPLS[[i]][[5+k]])[[1]])
    ## Get the PC scores
    pc_scores[[i]][[k]] <- as.data.frame(all_data[[i]][[k]]$ordination$x[,1])

    ## Get the centroid sizes
    centroid_size[[i]][[k]] <- as.data.frame(all_data[[i]][[k]]$procrustes$Csize)  
    
    ## Combine the pls data
    pls_pca[[i]][[k]] <- merge(TBPLS_scores[[i]][[k]], pc_scores[[i]][[k]],
                               by = "row.names", all.x = TRUE)
    ## Renaming the rows
    row.names(pls_pca[[i]][[k]]) <- pls_pca[[i]][[k]]$Row.names

    ## Combined the merged PLS
    merged_pls[[i]][[k]] <- merge(pls_pca[[i]][[k]], centroid_size[[i]][[k]],
                                  by = "row.names", all.x = TRUE)
    ## Renaming the rows
    row.names(merged_pls[[i]][[k]]) <- merged_pls[[i]][[k]]$Row.names

    ## Remove the two first columns
    merged_pls[[i]][[k]] <- merged_pls[[i]][[k]][,-(1:2)]
    ## Renaming the columns
    colnames(merged_pls[[i]][[k]]) <- c("PLS_Scores", "PC1","Csize")
  }

  ## Add the species names
  names(merged_pls)[[i]] <- names(all_data)[[i]]
  
  ## Add the dataset name (cranium/mandible)
  names(merged_pls[[i]]) <- names(all_data[[1]])
}
```

# Run the linear models of PLS scores with PC1 scores and centroid size

```{r, message = FALSE, warning = FALSE}
## Placeholder lists
pca_pls_cor <- c_size_pls_cor <- list()
 
##corelating the pls scores with PC1 scores and centroid sizes for each species and mandible/cranium
for (i in 1:length(merged_pls)){
    ## Placeholders
    pca_pls_cor[[i]] <- c_size_pls_cor[[i]] <- list()
    
    ##this runs the lm analyses for [1] PC scores and [2] Csize
    for (k in 1: length(merged_pls[[1]])){
        pca_pls_cor[[i]][[k]] <- list()
        ##
        pca_pls_cor[[i]][[k]][[1]] <- cor.test(
                                          merged_pls[[i]][[k]]$PLS_Scores , merged_pls[[i]][[k]]$PC1
                                            )
        ##
        pca_pls_cor[[i]][[k]][[2]] <- cor.test(
                                        merged_pls[[i]][[k]]$PLS_Scores , merged_pls[[i]][[k]]$Csize
                                            )
       
    }
    ## Renaming the list elements
    names(pca_pls_cor)[[i]] <- names(merged_pls)[[i]]
    names(pca_pls_cor[[i]]) <- names(merged_pls[[1]])
    names(pca_pls_cor[[i]][[k]]) <- c("pls_pca", "PLS_Csize")
}

```

## SHNW cranial PLS scores are the only ones with no significant correlation of PLS and PC1 scores. Therefore testing this correlation with PC2 as well

```{r}
#Turn cranial PLS scores and PC scores into a data frame for merging
SHNW_cranial_scores <- as.data.frame(SHNWTBPLS$XScores[c(1:19)],
                                           row.names = dimnames(SHNWTBPLS$XScores)[[1]])
SHNWPC2 <- as.data.frame(SHNW$cranium$ordination$x[,2])

#Merge data frames
SHNW_PLS_PC2<-merge(SHNW_cranial_scores,SHNWPC2,by="row.names", all.x=TRUE )
colnames(SHNW_PLS_PC2)=c("Species","PLS_Scores", "PC2")

#run analysis
cor.test(SHNW_PLS_PC2$PLS_Scores, SHNW_PLS_PC2$PC2)

```


