---
title: "Standard analyses"
author: "Vera Weisbecker"
bibliography: references.bib
date: "`r Sys.Date()`"
output:
  html_document:
    fig_width: 12
    fig_height: 6
---

```{r, message = FALSE, warning = FALSE}
## Loading the libraries (and installing if necessary)
if(!require(knitr)) install.packages("knitr"); library(knitr)
if(!require(xtable)) install.packages("xtable"); library(xtable)
if(!require(geomorph)) install.packages("geomorph"); library(geomorph)
if(!require(dispRity)) install.packages("dispRity");library(dispRity)
source("../Functions/utilities.R")
set.seed(42)

## Loading and renaming the land_data objects
load("../Data/Processed/wombat.Rda")
AllWombats <- land_data
load("../Data/Processed/wombat_krefftii.Rda")
NHNW <- land_data
load("../Data/Processed/wombat_latifrons.Rda")
SHNW <- land_data
load("../Data/Processed/wombat_ursinus.Rda")
CW <- land_data

## Read the classifiers
cranium <- read.csv("../Data/Raw/classifier_cranium_wombat.csv")
mandible <- read.csv("../Data/Raw/classifier_mandible_wombat.csv")
## Remove juvenile from mandibles
nhnw_b34 <- which(mandible[,1] == "NHNW_B34")
mandible <- mandible[-nhnw_b34, ]

##brief check that the landmarks were pulled out correctly - number of landmarks and plotting of landmarks
  ### fixed landmarks only: 65 fixed landmarks for cranium, 35 fixed landmarks for mandible
 length(dimnames(mshape(AllWombats$cranium$procrustes_no_patches$coords))[[1]]); plot3d(mshape(AllWombats$cranium$procrustes_no_patches$coords),asp=FALSE)
  
 length( dimnames(mshape(AllWombats$mandible$procrustes_no_patches$coords))[[1]])
 plot3d(mshape(AllWombats$mandible$procrustes_no_patches$coords),asp=FALSE)
  
   ### fixed landmarks and curves only: 261 curve landmarks for cranium, 142 fixed landmarks for mandible
  length(dimnames(mshape(AllWombats$cranium$procrustes_no_patches_no_curves$coords))[[1]])-65; plot3d(mshape(AllWombats$cranium$procrustes_no_patches$coords),asp=FALSE)
  
 length( dimnames(mshape(AllWombats$mandible$procrustes_no_patches_no_curves$coords))[[1]])-35;
 plot3d(mshape(AllWombats$mandible$procrustes_no_patches$coords),asp=FALSE)
  

```

#Separating the classifier file by species

```{r, message = FALSE, warning = FALSE}
## Creating the vectors holding the species names
species <- c("CW", "SHNW", "NHNW")
sp_holders <-c("Common wombat", "SHN wombat", "NHN wombat")

## Assigning the right class to each species (for the cranium)
for (i in 1:length(unique(cranium$Species))) {
    ## Extracting the class for each species   
    class <- cranium[cranium$Species == sp_holders[i],]
    ## Preparing the sp name for giving names to outputs
    sp_name <- paste(species[i], deparse(substitute(cranium)), "class", sep = "_")
    ## Assigning the class to the names
    assign(sp_name, class)
}

## Assigning the right class to each species (for the mandible)
for (i in 1:length(unique(mandible$Species))) {  
    ## Extracting the class for each species   
    class <- mandible[mandible$Species == sp_holders[i],]
    ## Preparing the sp name for giving names to outputs
    sp_name <- paste(species[i], deparse(substitute(mandible)), "class", sep = "_")
    ## Assigning the class to the names
    assign(sp_name, class)
}
```

#How much % variation is explained by PC1?

```{r}

plotTangentSpace(NHNW$cranium$procrustes_no_patches$coords)
plotTangentSpace(NHNW$mandible$procrustes_no_patches$coords)

plotTangentSpace(SHNW$cranium$procrustes_no_patches$coords)
plotTangentSpace(SHNW$mandible$procrustes_no_patches$coords)

plotTangentSpace(CW$cranium$procrustes_no_patches$coords)
plotTangentSpace(CW$mandible$procrustes_no_patches$coords)



```



# disparity comparison between wombat species

```{r, message = FALSE, warning = FALSE}
## Create gdfs with species for the data frames
cranium_gdf  <- geomorph.data.frame(coords = AllWombats$cranium$procrustes_no_patches$coords,
                                    Csize = AllWombats$cranium$procrustes_no_patches$Csize,
                                    species = as.factor(cranium$Species))
mandible_gdf <- geomorph.data.frame(coords = AllWombats$mandible$procrustes_no_patches$coords,
                                    Csize = AllWombats$mandible$procrustes_no_patches$Csize,
                                    species = as.factor(mandible$Species))


## Running the analyses on the gdf objects (Disparity within groups compared to that group's mean)
cranium_disparity <- morphol.disparity(coords ~ species, groups = ~species, iter = 1000, seed = 42,
                                       data = cranium_gdf, print.progress = FALSE)

## Disparity within groups compared to that group's mean
mandible_disparity <- morphol.disparity(coords ~ species, groups = ~species, iter = 1000, seed = 42,
                                        data = mandible_gdf, print.progress = FALSE)
```

And we can display the results as follows:

```{r, echo = FALSE}
## Cranium disparity values
kable(data.frame("Procrustes variance" = cranium_disparity[[1]]), digits = 5, caption = "Cranium disparity (procrustes variance) per groups (only fixed landmarks).")
## Cranium disparity differences
kable(cranium_disparity[[2]], digits = 5, caption = "Cranium pairwise absolute differences between variances between groups (only fixed landmarks).")
## Cranium disparity differences p
kable(cranium_disparity[[3]], digits = 5, caption = "Cranium pairwise absolute differences between variances between groups (p-values - only fixed landmarks).")

## Mandible disparity values
kable(data.frame("Procrustes variance" = mandible_disparity[[1]]), digits = 5, caption = "Mandible disparity (procrustes variance) per groups (only fixed landmarks).")
## Mandible disparity differences
kable(mandible_disparity[[2]], digits = 5, caption = "Mandible pairwise absolute differences between variances between groups (only fixed landmarks).")
## Mandible disparity differences p
kable(mandible_disparity[[3]], digits = 5, caption = "Mandible pairwise absolute differences between variances between groups (p-values - only fixed landmarks).")
```

#Are wombats sexually dimorphic?


## Make the sex subsets

The following analysis don't contain the northern hairy nose wombats because there is no variation in sex assignment in the dataset.

```{r, message = FALSE, warning = FALSE}
## Making the classifiers list
all_data <- list("SHNW" = SHNW, "CW" = CW)
all_classifiers <- list("SHNW" = list("cranium" = SHNW_cranium_class,
                                      "mandible" = SHNW_mandible_class),
                          "CW" = list("cranium" = CW_cranium_class,
                                      "mandible" = CW_mandible_class))
```

Creating the sex subsets and the centroid size per sex subsets:

```{r}
## Results placeholder
coords_sex_output <- c_size_sex_output <- list()

## Getting the coordinates for specimens with Sex
for (i in 1:length(all_data)){
    ## Result placeholder
    list() ->
        coords_sex_output[[i]] ->
        c_size_sex_output[[i]]

    ##Â Getting the coordinates for each classifier
    for (k in 1:length(all_classifiers[[i]])){
        
        ## Sex coordinates
        coords_sex_output[[i]][[k]] <- all_data[[i]][[k]]$procrustes_no_patches$coords[,,
                                       !is.na(all_classifiers[[i]][[k]]$Sex)]
        ## Sex centroid size
        c_size_sex_output[[i]][[k]] <- all_data[[i]][[k]]$procrustes_no_patches$Csize[
                                       !is.na(all_classifiers[[i]][[k]]$Sex)]

    }
    ## Naming the element
    names(all_data[[i]]) ->
      names(coords_sex_output[[i]]) ->
      names(c_size_sex_output[[i]])
}

names(all_data) ->
  names(coords_sex_output) ->
  names(c_size_sex_output)  
```

Creating the coordinates subsets and the centroid size subsets for sex:

```{r}
## Results placeholder
sex_gdfs <- list()

## Creating the geomorph dataframe for both species and datasets
for (i in 1:length(coords_sex_output)){
    ## Results placeholder
    sex_gdfs[[i]] <- list()
    
    for (k in 1:length(all_classifiers[[i]])){
      
      ## Sex data frame object
      sex_gdfs[[i]][[k]] <- geomorph.data.frame(
                          coords = coords_sex_output[[i]][[k]],
                          Csize = c_size_sex_output[[i]][[k]],
                          Sex = all_classifiers[[i]][[k]]$Sex[!is.na(all_classifiers[[i]][[k]]$Sex)]
                          )
  }
  ## Naming the elements
  names(sex_gdfs[[i]]) <- names(all_data[[i]])
}
names(sex_gdfs) <- names(all_data) 

## Making sure the dimension names are correct;requires testthat
#expect_equal(match(dimnames(geo_gdfs$SHNW$cranium$coords)[[3]],names(geo_gdfs$SHNW$cranium$Csize)),
#             seq(1:length(names(geo_gdfs$SHNW$cranium$Csize))))
```


## Running procD.lm analyses

We can run the three following linear models for each species and each datasets:

 * coordinates ~ sex + centroid size
 * centroid size ~ sex

```{r, message = FALSE, warning = FALSE}
## Results placeholder
coord_sex_lm <- csize_sex_lm <- list()

## Running all the linear models
for (i in 1:length(sex_gdfs)){
    ## Results placeholder
   coord_sex_lm[[i]] <- csize_sex_lm[[i]] <- list()
        
    for (k in 1:length(sex_gdfs[[2]])){
        
        ## Running the coordinates ~ sex + centroid size models
        coord_sex_lm[[i]][[k]] <- procD.lm(coords ~ Sex , print.progress = FALSE,
                                           data = sex_gdfs[[i]][[k]], iter = 1000)
        ## Running the centroid size ~ sex models
        csize_sex_lm[[i]][[k]] <- procD.lm(Csize ~ Sex, print.progress = FALSE,
                                           data = sex_gdfs[[i]][[k]], iter = 1000)
    }
    ## Renaming the models 
    names(coord_sex_lm[[i]]) -> names(csize_sex_lm[[i]])
  }

names(coord_sex_lm) <- names(csize_sex_lm) 
```


And we can display them as follows:

```{r, echo = FALSE}

## coordinates ~ sex + centroid size
kable(coord_sex_lm[[1]][[1]]$aov.table, digits = 3,
    caption = "Procrustes (only fixed landmarks) linear model for coordinates as a function of sex and centroid size for the southern hairy nose wombat's cranium.")
kable(coord_sex_lm[[1]][[2]]$aov.table, digits = 3,
    caption = "Procrustes (only fixed landmarks) linear model for coordinates as a function of sex and centroid size for the southern hairy nose wombat's mandible.")
kable(coord_sex_lm[[2]][[1]]$aov.table, digits = 3,
    caption = "Procrustes (only fixed landmarks) linear model for coordinates as a function of sex and centroid size for the common wombat's cranium.")
kable(coord_sex_lm[[2]][[2]]$aov.table, digits = 3,
    caption = "Procrustes (only fixed landmarks) linear model for coordinates as a function of sex and centroid size for the common wombat's mandible.")

## centroid size ~ sex 
kable(csize_sex_lm[[1]][[1]]$aov.table, digits = 3,
    caption = "Procrustes (only fixed landmarks) linear model for centroid size as a function of sex for the southern hairy nose wombat's cranium.")
kable(csize_sex_lm[[1]][[2]]$aov.table, digits = 3,
    caption = "Procrustes (only fixed landmarks) linear model for centroid size as a function of sex for the southern hairy nose wombat's mandible.")
kable(csize_sex_lm[[2]][[1]]$aov.table, digits = 3,
    caption = "Procrustes (only fixed landmarks) linear model for centroid size as a function of sex for the common wombat's cranium.")
kable(csize_sex_lm[[2]][[2]]$aov.table, digits = 3,
    caption = "Procrustes (only fixed landmarks) linear model for centroid size as a function of sex for the common wombat's mandible.")
```


# Allometry Analyses

We can run the allometry tests using the `handle.procD.formula` wrapper function:

```{r, message = FALSE, warning = FALSE, result = 'hide'}
## List of all the data
all_data <- list("NHNW" = NHNW, "SHNW" = SHNW, "CW" = CW)

## Placeholder for the results
allom_output <- list()

## Running the allometry tests
for (i in 1:length(all_data)){
    for (k in 1:length(all_data[[1]])){
        ## Result placeholder
        allom_output[i][[k]] <- list()

        ## Procrustes allometry test (wrapper around the procD.allometry function)
        allom_output[[i]][[k]] <- handle.procD.formula(coords ~ Csize,
                                                 all_data[[i]][[k]]$procrustes_no_patches,
                                                 procD.fun = procD.lm, iter = 1000,
                                                 print.progress = FALSE)
    }
}

## Naming the results
names(allom_output) <- names(all_data)

## Naming the partitions
for(i in 1:length(all_data)) {
    names(allom_output[[i]]) <- names(all_data[[i]])
}
```

And we can display them as follows:

```{r, echo = FALSE}
kable(allom_output[[1]][[1]]$aov.table, digits = 3,
    caption = "Procrustes ANOVA for size (allometry) for the northern hairy nose wombat's cranium (only fixed landmarks).")
kable(allom_output[[1]][[2]]$aov.table, digits = 3,
   caption = "Procrustes ANOVA for size (allometry) for the northern hairy nose wombat's mandible (only fixed landmarks).")
kable(allom_output[[2]][[1]]$aov.table, digits = 3,
    caption = "Procrustes ANOVA for size (allometry) for the southern hairy nose wombat's cranium (only fixed landmarks).")
kable(allom_output[[2]][[2]]$aov.table, digits = 3,
   caption = "Procrustes ANOVA for size (allometry) for the southern hairy nose wombat's mandible (only fixed landmarks).")
kable(allom_output[[3]][[1]]$aov.table, digits = 3,
              caption = "Procrustes ANOVA for size (allometry) for the common wombat's cranium (only fixed landmarks).")
kable(allom_output[[3]][[2]]$aov.table, digits = 3,
              caption = "Procrustes ANOVA for size (allometry) for the common wombat's mandible (only fixed landmarks).")
```

#Allometry between species

```{r}
#Cranium

##turn into gdf frame 
gdf_cranium <- geomorph.data.frame(coords = AllWombats$cranium$procrustes_no_patches$coords, 
                           Csize = AllWombats$cranium$procrustes_no_patches$Csize)

##run analysis
AllomCranium <- procD.lm(coords~Csize, data = gdf_cranium, method = "RegScore")

#Mandible

##turn into gdf frame 
gdf_mandible <- geomorph.data.frame(coords = AllWombats$mandible$procrustes_no_patches$coords,
                           Csize = AllWombats$mandible$procrustes_no_patches$Csize)

##run analysis
AllomMandible <- procD.lm(coords~Csize, data = gdf_mandible, method = "RegScore")
```


# 2-Block Partial Least Squares analysis (2BPLS)

## Partition classifiers according to species

<!-- Full_classifier_Table_ is the tables of mandibular or cranial classifiers (e.g. read.csv("../Data/Raw/classifier_cranium_wombat.csv")); spp is a vector of species (e.g. c("CW", "SHNW", "NHNW")) -->


## Reduce coordinate datasets to only those that have mandibles and crania

```{r, message = FALSE, warning = FALSE}
## Data list
all_data <- list("NHNW" = NHNW, "SHNW" = SHNW, "CW" = CW)
## Classifiers list
all_classifiers <- list(
                    "NHNW" = list("cranium" = NHNW_cranium_class, "mandible" = NHNW_mandible_class),
                    "SHNW" = list("cranium" = SHNW_cranium_class, "mandible" = SHNW_mandible_class),
                    "CW" = list("cranium" = CW_cranium_class, "mandible" = CW_mandible_class)
                   )

## Checking whether the cranial and mandibular coordinates datasets match; see utilities for function
matched_coords <- reduce.check(all_data, all_classifiers, procrustes=4)
```

## Run 2BPLS

Depending on processor speed, this can take a while, so we recommend just loading the results which are saved in the `Results` folder.
The results can be run using the following code chunk:


```{r, message = FALSE, warning = FALSE, eval = TRUE}
## 2BPLS for the southern hairy nose wombats
SHNWTBPLS <- two.b.pls(matched_coords[[1]]$SHNW$cranium ,matched_coords[[1]]$SHNW$mandible,
                       iter = 1000) 
save(SHNWTBPLS, file = "../Data/Results/SHNWTBLS_nocurves_nopatches.Rda")

## 2BPLS for the northern hairy nose wombats
NHNWTBPLS <- two.b.pls(matched_coords[[1]]$NHNW$cranium ,matched_coords[[1]]$NHNW$mandible,
                       iter = 1000) 
save(NHNWTBPLS, file="../Data/Results/NHNWTBLS_nocurves_nopatches.Rda")

## 2BPLS for the common wombats
CWTBPLS <- two.b.pls(matched_coords[[1]]$CW$cranium ,matched_coords[[1]]$CW$mandible,
                     iter = 1000) 
save(CWTBPLS, file="../Data/Results/CWTBLS_nocurves_nopatches.Rda")
```

Or directly loaded using this one:
 
```{r}
## 2BPLS for the southern hairy nose wombats
load("../Data/Results/SHNWTBLS_nocurves_nopatches.Rda")
## 2BPLS for the northern hairy nose wombats
load("../Data/Results/CWTBLS_nocurves_nopatches.Rda")
## 2BPLS for the common wombats
load("../Data/Results/NHNWTBLS_nocurves_nopatches.Rda")
```

## Subsetting data to run correlations of PLS score with csize and allometry

```{r, message = FALSE, warning = FALSE}
## Getting the list of 2BPLS results
all_TBPLS <- list("NHNW" = NHNWTBPLS, "SHNW" = SHNWTBPLS, "CW" = CWTBPLS)

## Placeholder lists
pls_pca_c_size <- TBPLS_scores <- pc_scores <- centroid_size <- pls_pca <- merged_pls <- list()

## Creating the subsets
for (i in 1:length(all_data)){
 
  ## Placeholder lists
  list() -> 
      pls_pca_c_size [[i]] ->
      TBPLS_scores[[i]] ->
      pc_scores[[i]] -> 
      centroid_size[[i]] ->
      pls_pca [[i]] ->
      merged_pls[[i]]

  ## Create data frames for subsequent merging. 
  for (k in 1:length(all_data$NHNW)){

    ## TBPLS needs to contain i(species), an address for k (x/yscores as per cranium/mandible, and a
    ## designation of how long the first singular vector goes, i.e. the species number); the 5+k i
    ## the "address" of cranial/mandibular scores in the TBPLS data frames
    TBPLS_scores[[i]][[k]] <- as.data.frame(all_TBPLS[[i]][[5+k]]
                                            [c(1:length(dimnames(all_TBPLS[[i]][[5+k]])[[1]]))],
                                            row.names = dimnames(all_TBPLS[[i]][[5+k]])[[1]])
    ## Get the PC scores
    pc_scores[[i]][[k]] <- as.data.frame(all_data[[i]][[k]]$ordination$x[,1])

    ## Get the centroid sizes
    centroid_size[[i]][[k]] <- as.data.frame(all_data[[i]][[k]]$procrustes_no_patches$Csize)  
    
    ## Combine the pls data
    pls_pca[[i]][[k]] <- merge(TBPLS_scores[[i]][[k]], pc_scores[[i]][[k]],
                               by = "row.names", all.x = TRUE)
    ## Renaming the rows
    row.names(pls_pca[[i]][[k]]) <- pls_pca[[i]][[k]]$Row.names

    ## Combined the merged PLS
    merged_pls[[i]][[k]] <- merge(pls_pca[[i]][[k]], centroid_size[[i]][[k]],
                                  by = "row.names", all.x = TRUE)
    ## Renaming the rows
    row.names(merged_pls[[i]][[k]]) <- merged_pls[[i]][[k]]$Row.names

    ## Remove the two first columns
    merged_pls[[i]][[k]] <- merged_pls[[i]][[k]][,-(1:2)]
    ##Â Renaming the columns
    colnames(merged_pls[[i]][[k]]) <- c("PLS_Scores", "PC1","Csize")
  }

  ## Add the species names
  names(merged_pls)[[i]] <- names(all_data)[[i]]
  
  ## Add the dataset name (cranium/mandible)
  names(merged_pls[[i]]) <- names(all_data[[1]])
}
```

# Run the linear models of PLS scores with PC1 scores and centroid size

```{r, message = FALSE, warning = FALSE}
## Placeholder lists
pca_pls_cor <- c_size_pls_cor <- list()
 
##corelating the pls scores with PC1 scores and centroid sizes for each species and mandible/cranium
for (i in 1:length(merged_pls)){
    ##Placeholders
    pca_pls_cor[[i]] <- c_size_pls_cor[[i]] <- list()
    
    ##this runs the lm analyses for [1] PC scores and [2] Csize
    for (k in 1: length(merged_pls[[1]])){
        pca_pls_cor[[i]][[k]] <- list()
        ##
        pca_pls_cor[[i]][[k]][[1]] <- cor.test(
                                          merged_pls[[i]][[k]]$PLS_Scores , merged_pls[[i]][[k]]$PC1
                                            )
        ##
        pca_pls_cor[[i]][[k]][[2]] <- cor.test(
                                        merged_pls[[i]][[k]]$PLS_Scores , merged_pls[[i]][[k]]$Csize
                                            )
       
    }
    ## Renaming the list elements
    names(pca_pls_cor)[[i]] <- names(merged_pls)[[i]]
    names(pca_pls_cor[[i]]) <- names(merged_pls[[1]])
    names(pca_pls_cor[[i]][[k]]) <- c("pls_pca", "PLS_Csize")
}

```

## SHNW cranial PLS scores are the only ones with no significant correlation of PLS and PC1 scores. Therefore testing this correlation with PC2 as well

```{r}
#Turn cranial PLS scores and PC scores into a data frame for merging
SHNW_cranial_scores <- as.data.frame(SHNWTBPLS$XScores[c(1:19)],
                                           row.names = dimnames(SHNWTBPLS$XScores)[[1]])
SHNWPC2 <- as.data.frame(SHNW$cranium$ordination$x[,2])

#Merge data frames
SHNW_PLS_PC2<-merge(SHNW_cranial_scores,SHNWPC2,by="row.names", all.x=TRUE )
colnames(SHNW_PLS_PC2)=c("Species","PLS_Scores", "PC2")

#run analysis
cor.test(SHNW_PLS_PC2$PLS_Scores, SHNW_PLS_PC2$PC2)

```


