---
title: "Heat map plot of changes"
author: "Thomas Guillerme"
date: "`r Sys.Date()`"
output:
  html_document:
    fig_width: 12
    fig_height: 6
---

Some fancy plots for the maximum landmark variation.

## Plots 

### Loading the functions

```{r, message = FALSE, warning = FALSE}
## Loading the libraries (and installing if necessary)
if(!require(devtools)) install.packages("devtools"); library(devtools)
if(!require(geomorph)) install.packages("geomorph"); library(geomorph)
if(!require(dispRity)) install.packages("dispRity"); library(dispRity)
if(!require(landvR)) install_github("TGuillerme/landvR") ; library(landvR)
set.seed(42)
```


### Loading the data

```{r}
## Loading a dataset

#loading the below loads objects called land_data, NOT "wombat"
load("../Data/Processed/wombat.Rda")
AllWombats<-land_data
load("../Data/Processed/wombat_lasiorhinus.Rda")
HN<-land_data
load("../Data/Processed/wombat_krefftii.Rda")
NHNW<-land_data
load("../Data/Processed/wombat_latifrons.Rda")
SHNW<-land_data
load("../Data/Processed/wombat_ursinus.Rda")
CW<-land_data
source("../Functions/utilities.R")

```

The "maximum" and "minimum" specimen in procrustes space

```{r}
## Procrustes variation ranges
cranium=SHNW$cranium
variation <- variation.range(cranium$procrustes, return.ID = TRUE)
specimens_min_max <- variation$min.max
procrustes_var <- variation$range[,1]
```

```{r, eval = FALSE}
gridPar = gridPar(pt.bg = "white", pt.size = 0.5)
open3d()
procrustes.var.plot(cranium$procrustes$coords[, , specimens_min_max[2]],
                     cranium$procrustes$coords[, , specimens_min_max[1]],
                     col = heat.colors, pt.size = 0.8, col.val = procrustes_var)
```

Specimens along PC extremes (i.e. not hypothetical specimens, as per warning message)

```{r}
##  variation ranges
cranium=CW$cranium
variation <- variation.range(cranium$procrustes, return.ID = TRUE, axis=1, ordination=cranium$ordination )
variation_hyp <- variation.range(cranium$procrustes, axis=1, ordination=cranium$ordination )

specimens_min_max <- variation$min.max
procrustes_var <- variation$range[,1]
```

```{r, eval = FALSE}
gridPar = gridPar(pt.bg = "white", pt.size = 0.5)
open3d()
procrustes.var.plot(cranium$procrustes$coords[, , specimens_min_max[2]],
                     cranium$procrustes$coords[, , specimens_min_max[1]],
                     col = heat.colors, pt.size = 0.8, col.val =  variation$range[,1])
```


```{r}
## Comparing the coordinates of both hypothetical and extreme specimens.
species_dataset <- AllWombats$cranium

## Getting the true specimens that are closest to the hypothetical ones
variation <- variation.range(species_dataset$procrustes, return.ID = TRUE, axis = 1,
                             ordination = species_dataset$ordination)
true_specimens <- variation$min.max

## Getting the hypothetical specimens (as output by plotTangentSpace)
hypothetical_spec <- plotTangentSpace(species_dataset$procrustes$coords, verbose = FALSE)

## Isolating the pairs of specimens
hypothetical_spec_min <- hypothetical_spec$pc.shapes[[1]]
hypothetical_spec_max <- hypothetical_spec$pc.shapes[[2]]
true_specimens_min <- species_dataset$procrustes$coords[, , true_specimens[1]]
true_specimens_max <- species_dataset$procrustes$coords[, , true_specimens[2]]

## Comparing the max and mins (they are supposed to be different!)
all(hypothetical_spec_min == hypothetical_spec_max) # FALSE (they are different)
all(true_specimens_min == true_specimens_max) # FALSE

## Comparing the true and the hypothetical (the are supposed to be different)
all(hypothetical_spec_min == true_specimens_min) # FALSE
all(hypothetical_spec_max == true_specimens_max) # FALSE

## Are any of the coordinates the same?
any(hypothetical_spec_min == true_specimens_min) # FALSE
any(hypothetical_spec_max == true_specimens_max) # FALSE

## Is it a rounding problem (how many are similar per rounding)
## This function returns the percentage of coordinates that are matching
compare.rounding <- function(digits, x, y) {
   rounding_equals <- length(which(round(x, digits) == round(y, digits)))
   return(rounding_equals/length(x)*100)
}
differences_min <- sapply(c(1:7), compare.rounding, hypothetical_spec_min, true_specimens_min)
differences_max <- sapply(c(1:7), compare.rounding, hypothetical_spec_max, true_specimens_max)

## Plotting these results
plot(NULL, xlab = "Precision (digits)", ylab = "Percentage of coordinates matching",
           ylim = c(0,100), xlim = c(1,7))
legend("topright", lty = 1, col = c("purple", "pink"),
       legend = c("hypothetical vs. max proj. (min)", "hypothetical vs. max proj. (max)"))
lines(differences_min, col = "purple")
lines(differences_max, col = "pink")

```


<!-- ``` -->

#Doing this using hypothetical PC min/max landmarks for the figures in the manuscript. See utilities file for details. Also be aware that the warning message that the returned specimen does not correspond to the PCmin/max specimen but the nearest observed one is INCORRECT in this case becasue we are using the hypothetical configurations as specimens.

```{r, eval = FALSE}
## Craniums
## All wombats
heatplot.PCs(AllWombats$cranium, minfirst=FALSE, 1 )
## Lasiorhinus
heatplot.PCs(HN$cranium, minfirst=TRUE ,1 )
## Vombatus
heatplot.PCs(CW$cranium, minfirst=FALSE,1)
## Lasihorhinus latifrons
heatplot.PCs(SHNW$cranium, minfirst=TRUE,1)
## Lasihorhinus krefftii
heatplot.PCs(NHNW$cranium, minfirst=TRUE,1)

## Mandibles
## All wombats
heatplot.PCs(AllWombats$mandible, minfirst=FALSE, 1)
## Lasihorhinus
heatplot.PCs(HN$mandible, minfirst=TRUE, 1 )
## Vombatus
heatplot.PCs(CW$mandible, minfirst=FALSE,1)
## Lasihorhinus latifrons
heatplot.PCs(SHNW$mandible, minfirst=TRUE,1)
## Lasihorhinus krefftii
heatplot.PCs(NHNW$mandible, minfirst=TRUE, 1)
```

