---
title: "Heat map plot of changes"
author: "Thomas Guillerme"
date: "`r Sys.Date()`"
output:
  html_document:
    fig_width: 12
    fig_height: 6
---

Some fancy plots for the maximum landmark variation.

## Plots 

### Loading the functions

```{r, message = FALSE, warning = FALSE}
## Loading the libraries (and installing if necessary)
if(!require(devtools)) install.packages("devtools"); library(devtools)
if(!require(geomorph)) install.packages("geomorph"); library(geomorph)
if(!require(dispRity)) install.packages("dispRity"); library(dispRity)
if(!require(landvR)) install_github("TGuillerme/landvR") ; library(landvR)
set.seed(42)
```


### Loading the data

```{r}
## Loading a dataset

#loading the below loads objects called land_data, NOT "wombat"
load("../Data/Processed/wombat.Rda")
AllWombats<-land_data
load("../Data/Processed/wombat_lasiorhinus.Rda")
HN<-land_data
load("../Data/Processed/wombat_krefftii.Rda")
NHNW<-land_data
load("../Data/Processed/wombat_latifrons.Rda")
SHNW<-land_data
load("../Data/Processed/wombat_ursinus.Rda")
CW<-land_data
source("../Functions/utilities.R")

```

The "maximum" and "minimum" specimen in procrustes space

```{r}
## Procrustes variation ranges
target=CW$cranium
variation <- variation.range(target$procrustes, return.ID = TRUE)
specimens_min_max <- variation$min.max
procrustes_var <- variation$range[,1]
```

```{r, eval = FALSE}
gridPar = gridPar(pt.bg = "white", pt.size = 0.5)
open3d()
procrustes.var.plot(target$procrustes$coords[, , specimens_min_max[1]],
                     target$procrustes$coords[, , specimens_min_max[2]],
                     col = heat.colors, pt.size = 0.8, col.val = procrustes_var, main = "GPA var")
```

Specimens along PC extremes (i.e. not hypothetical specimens, as per warning message)

```{r}
##  variation ranges
target=CW$cranium
#HIgest and lowest Pc1 score - specimen closest to PC1 extremes
variation_real <- variation.range(target$procrustes, axis=1, ordination=target$ordination, return.ID = TRUE)

minmax <- variation_real$min.max

#measuring the difference between the two sepcimens selected in variation.real
variation <- coordinates.difference(target$procrustes$coords[,,minmax[1]], target$procrustes$coords[,,minmax[2]], type = "spherical")

open3d()
#Use "variation" to colour the plot exactly according to vector length
procrustes.var.plot(target$procrustes$coords[, , minmax[2]],
                     target$procrustes$coords[, , minmax[1]],
                     col = heat.colors, pt.size = 0.8, col.val = variation[[1]][,1],
                      main="Min-max specimens closes to PC1 extremes")

```

```{r, eval = FALSE}
gridPar = gridPar(pt.bg = "white", pt.size = 0.5)
open3d()
procrustes.var.plot(target$procrustes$coords[, , specimens_min_max[2]],
                     target$procrustes$coords[, , specimens_min_max[1]],
                     col = heat.colors, pt.size = 0.8, col.val =  variation$range[,1],
                      main="procrustes.var.plot based actual specimens")
```


#Doing this using hypothetical PC min/max landmarks for the figures in the manuscript. See utilities file for details. Also be aware that the warning message that the returned specimen does not correspond to the PCmin/max specimen but the nearest observed one is INCORRECT in this case becasue we are using the hypothetical configurations as specimens.

```{r, eval = FALSE}
## Craniums
## All wombats
heatplot.PCs(AllWombats$cranium, minfirst=FALSE, 1 )
## Lasiorhinus
heatplot.PCs(HN$cranium, minfirst=TRUE ,1 )
## Vombatus
heatplot.PCs(CW$cranium, minfirst=FALSE,1, main="hypothetical CW PC1")
## Lasihorhinus latifrons
heatplot.PCs(SHNW$cranium, minfirst=TRUE,1)
## Lasihorhinus krefftii
heatplot.PCs(NHNW$cranium, minfirst=TRUE,1)

## Mandibles
## All wombats
heatplot.PCs(AllWombats$mandible, minfirst=FALSE, 1)
## Lasihorhinus
heatplot.PCs(HN$mandible, minfirst=TRUE, 1 )
## Vombatus
heatplot.PCs(CW$mandible, minfirst=FALSE,1)
## Lasihorhinus latifrons
heatplot.PCs(SHNW$mandible, minfirst=TRUE,1)
## Lasihorhinus krefftii
heatplot.PCs(NHNW$mandible, minfirst=TRUE, 1)
```

#In case there are any doubts that the heatplot and "regular" PlotTangentSpace/plotRefToTArget show the same, below is a comparison
```{r}

gridpar=gridPar(pt.size=0.3)

#Run conventional PlotRefToTarget
CWGeomorph=plotTangentSpace(CW$cranium$procrustes$coords)

#Find min/max specimen
which(CWGeomorph$pc.scores[,1] ==  max(CWGeomorph$pc.scores[,1]))#CW_M22233, number 3
which(CWGeomorph$pc.scores[,1] ==  min(CWGeomorph$pc.scores[,1]))#CW_M22233, number 14

#also no. 3 and 14 in the coordinates
dimnames(CW$cranium$procrustes$coords)[[3]]

#Plot hypothetical configurations 
open3d()
plotRefToTarget(CWGeomorph$pc.shapes$PC1max, CWGeomorph$pc.shapes$PC1min, method="vector",gridPar=gridpar, main="min/max hypothetical specimens on PC1 as determined through PlotTangentSpace")

#Plot actual specimens - they are different!

open3d()
plotRefToTarget( CW$cranium$procrustes$coords[,,3], CW$cranium$procrustes$coords[,,14], method="vector", gridPar=gridpar, main="min/max actual specimens of CW PC1 as determined through plotTangentSpace")


#hypothetical PCs "the old fashioned way" are comparable to the heat plots and running the above code also means that 
open3d()
heatplot.PCs(CW$cranium, minfirst=FALSE,1)

```

