---
title: "Figures/supplementary figures for wombat shape variation paper, except heatplots"
author: "Vera Weisbecker"
date: "`r Sys.Date()`"
output:
  html_document:
    fig_width: 12
    fig_height: 6
---


# Loading the data

```{r, message = FALSE, warning = FALSE}
## Loading the libraries (and installing if necessary)

if(!require(devtools)) install.packages("devtools"); library(devtools)
if(!require(geomorph)) install.packages("geomorph"); library(geomorph)
if(!require(landvR)) install_github("TGuillerme/landvR") ; library(landvR)
if(!require(stringr)) install.packages("stringr"); library(stringr)


library(ggplot2)

# Install and load ggConvexHull
#devtools::install_github("cmartin/ggConvexHull")

library(ggConvexHull)

source("../Functions/utilities.R")
set.seed(42)



# load the data
#for hairy-nosed wombats (needed below for PCA)
load("../Data/Processed/wombat_lasiorhinus.Rda")
HN <- land_data

##the below loads the Wombat.Rda from the processed data (i.e. partitions, procrustes data, and ordination results); but the object is NOT called the file name called under "load", it is "land_data"
load("../Data/Processed/wombat.Rda")
rawdata_cranium_wombat <- read.csv ("../Data/Raw/landmarkdata_cranium_wombat.csv")
rawdata_mandible_wombat <- read.csv ("../Data/Raw/landmarkdata_mandible_wombat.csv")
## Remove juvenile
nhnw_b34 <- which(colnames(rawdata_mandible_wombat) == "NHNW_B34")
rawdata_mandible_wombat <- rawdata_mandible_wombat[,-nhnw_b34]

# load allometry residuals

load("../Data/Processed/Allometry_residuals.rda")

#read classifiers
cranium <- read.csv("../Data/Raw/classifier_cranium_wombat.csv")
mandible <- read.csv("../Data/Raw/classifier_mandible_wombat.csv")
## Remove juvenile
nhnw_b34 <- which(mandible[,1] == "NHNW_B34")
mandible <- mandible[-nhnw_b34, ]




```


# PCA plot for Figure 1 

```{r, message = FALSE, warning = FALSE}

## Loading the data for the PCA plots
load("../Data/Processed/wombat.Rda")
classifier_cranium <- cranium
classifier_mandible <- mandible

##loading hues
gg.color.hue <- function(n) {
    hues = seq(15, 375, length = n + 1)
    hcl(h = hues, l = 65, c = 100)[1:n]
}

## PCA plots - Figure 1


pdf("../Manuscript/Figures/Figure1_PCA.pdf", height = 10, width = 12)
par(mfrow = c(2,2))
plot.pca(land_data$cranium$ordination, classifier_cranium$Species, cex.axis=1.2  );legend("bottomleft", c("Common", "Southern HN", "Northern HN"), pt.bg=(gg.color.hue(3)),pch=21, bty="n", cex=1.2); legend ("top", "All Species - Cranium", bty="n", cex=1.5)
plot.pca(land_data$mandible$ordination, classifier_mandible$Species,cex.axis=1.2 ); legend ("top", "All Species - Mandible", bty="n", cex=1.5)


## Loading the data for the within-species PCA plots
load("../Data/Processed/wombat_lasiorhinus.Rda")

classifier_cranium_Lasio= subset(classifier_cranium, classifier_cranium$Species != "Common wombat" )
classifier_mandible_Lasio= subset(classifier_mandible, classifier_mandible$Species != "Common wombat" )


plot.pca(HN$cranium$ordination, classifier_cranium_Lasio$Species, cex.axis=1.5 );  legend ("top", "Hairy-nosed - Cranium", bty="n", cex=1.5)
plot.pca(HN$mandible$ordination, classifier_mandible_Lasio$Species, cex.axis=1.5 );  legend ("top", "Hairy-nosed - Mandible", bty="n", cex=1.5)

dev.off()


```


#the same plot with allometry residuals (residuals of the regression of shape against centroid size)

```{r, message = FALSE, warning = FALSE}

load("../Data/Processed/Allometry_residuals.rda")


#make ordination as required for plot.pca (in utilities)
array_2d_cran <- geomorph::two.d.array(Allom_plus_consensus_cran)

ordination_cran <- prcomp(array_2d_cran, center = TRUE, scale. = FALSE, retx = TRUE,
                                              tol = NULL)
# Mandible, all species

array_2d_mand <- geomorph::two.d.array(Allom_plus_consensus_mand)

ordination_mand <- prcomp(array_2d_mand, center = TRUE, scale. = FALSE, retx = TRUE,
                                              tol = NULL)

#Cranium, Lasiorhinus
array_2d_cran_las <- geomorph::two.d.array(Allom_plus_consensus_cran_Lasio)

ordination_cran_las <- prcomp(array_2d_cran_las, center = TRUE, scale. = FALSE, retx = TRUE,
                                              tol = NULL)
# Mandible, Lasiorhinus

array_2d_mand_las <- geomorph::two.d.array(Allom_plus_consensus_mand_Lasio)

ordination_mand_las <- prcomp(array_2d_mand_las, center = TRUE, scale. = FALSE, retx = TRUE,
                                              tol = NULL)



classifier_cranium <- cranium
classifier_mandible <- mandible

##loading hues
gg.color.hue <- function(n) {
    hues = seq(15, 375, length = n + 1)
    hcl(h = hues, l = 65, c = 100)[1:n]
}

## PCA plots - supplementary fig. 2


pdf("../Manuscript/Figures/SuppX_Residual_PCA.pdf", height = 10, width = 12)
#plot.new()
par(mfrow = c(2,2))
plot.pca(ordination_cran, classifier_cranium$Species, cex.axis=1.2 );legend("bottomleft", c("Common", "Southern HN", "Northern HN"), pt.bg=(gg.color.hue(3)),pch=21, bty="n", cex=1.2 ); legend ("top", "All Species - Cranium", bty="n", cex=1.5)

plot.pca(ordination_mand, classifier_mandible$Species, cex.axis=1.2); legend ("top", "All Species - Mandible", bty="n", cex=1.5)


## Loading the data for the within-species PCA plots
load("../Data/Processed/wombat_latifrons.Rda")
classifier_cranium_Lasio= subset(classifier_cranium, classifier_cranium$Species != "Common wombat" )
classifier_mandible_Lasio= subset(classifier_mandible, classifier_mandible$Species != "Common wombat" )


plot.pca(ordination_cran_las, classifier_cranium_Lasio$Species, cex.axis=1.2);  legend ("top", "Hairy-nosed - Cranium", bty="n", cex=1.5)
plot.pca(ordination_mand_las, classifier_mandible_Lasio$Species, cex.axis=1.2);  legend ("top", "Hairy-nosed - Mandible", bty="n", cex=1.5 )

dev.off()


```





# Pull out landmark designations. These need to be raw data because the landmark names are in the raw data



```{r}


  ## Create vectors of patch points, semi-landmarks, and landmarks; this requires raw data because they contain the landmark designations.
  # LM point name metadata
  
#change "cranium" designations to "mandible" designations as needed



pt_names <-rawdata_mandible_wombat[, 1] # grabs first column of raw coords
foo2 <- str_sub(pt_names, 3, -1) # gets rid of X, Y, Z designations
u_pt_names <- unique(foo2)

patches <- str_detect(u_pt_names, "patch") # 

pat_num <- which(patches == TRUE)


sliders <- str_detect(u_pt_names, "CUR")
sli_num <- which(sliders == TRUE)

LM <- !(sliders | patches)
LM_num <- which(LM == TRUE)

#Assign each landmark the appropriate number; Fixed is 1, semis are 2, patches are 3

#using landmarkgroups as a file to fill in 
landmarkgroups_for_type=read.csv("../Data/Raw/landmarkgroups_mandible_wombat.csv")

landmark_LM_types<-landmarkgroups_for_type
for (i in 1:length(landmark_LM_types[,1])){
if(landmark_LM_types[i,1] %in% pat_num ==TRUE)
landmark_LM_types[i,2]<-3

else
  
if(landmark_LM_types[i,1] %in% sli_num ==TRUE)
landmark_LM_types[i,2]<-2 

else
  
if(landmark_LM_types[i,1] %in% LM_num ==TRUE)
landmark_LM_types[i,2]<-1   

}

write.csv(landmark_LM_types, "../Data/Raw/landmarktypes_mandible_wombat.csv")


``` 

#Visualize landmark types or landmark partitions without ply file

```{r, message = FALSE, warning = FALSE}
#separate land_data into the relevant components

#Define the partitions (needs to be done individually); this produces a csv file that can be manually edited (see comments below)
#LandmarkRegions=define.modules(land_data$cranium$procrustes$coords[,,5],2)
#write.csv(LandmarkRegions, file="../Data/Results/additional_snout.csv")

#Optional:
PartNames<-c("Zyg", "Sn", "Oc", "Po", "Rest")# This is not necessary for colouring the cranium AND needs to be separately accessed, but helps in identifying the partitionss

plot.partitions(land_data$cranium, PointSize = 0.001)
plot.partitions(land_data$mandible, PointSize = 0.001)

#Visualize partitions as colours
plot.partitions(land_data$cranium, PointSize = 0.001)

#optional - adding text helps with identifying landmarks if define.modules has not captured all landmarks of a partition. These numbers can then be manually edited in the LandmarkRegions file. 
#text3d(mshape(land_data$cranium$procrustes$coords), text=c(1:826), adj = 2, cex = 0.8)

```

#Visualize landmark types or landmark partitions superimposed on a surface using a ply file; more involved as it requires the use of raw data

```{r, eval = FALSE}

#Create a ply file of the mean shape by warping the mean specimen in the PCA to the mean shape; for this, non-procrustes superimposed coordinates are required. The mean shapes are used here because they are easy to register with the landmarks.

#Find mean specimen, determine mean shape to warp mean specimen to
findMeanSpec(land_data$cranium$procrustes$coords)
findMeanSpec(land_data$mandible$procrustes$coords)
mshape_cranium<-mshape(land_data$cranium$procrustes$coords)
mshape_mandible<-mshape(land_data$mandible$procrustes$coords)

#Load ply file for  specimens closest to mean
mean_specimen_mesh_cranium<-read.ply("../Data/Raw/NHNW_JM12487_Cranium.ply")
mean_specimen_mesh_mandible<-read.ply("../Data/Raw/SHNW NR1326 Mandible.ply")

#read raw data needed to register to ply files 
WomCrData <- read.csv("../Data/Raw/landmarkdata_cranium_wombat.csv", header = T, row.names = 1) 
WomCrData <- t(WomCrData) 
WomCrData <- WomCrData[complete.cases(WomCrData),] 
WomCrData <- arrayspecs(WomCrData, k=3, p=ncol(WomCrData)/3) 

WomMaData <- read.csv("../Data/Raw/landmarkdata_mandible_wombat.csv", header = T, row.names = 1) 
WomMaData <- t(WomMaData)
WomMaData <- WomMaData[complete.cases(WomMaData),] 
WomMaData <- arrayspecs(WomMaData, k=3, p=ncol(WomMaData)/3) 

# raw coordinates of the mean specimen to register to their respective ply files
mean_coords_cranium=WomCrData[,,38]
mean_coords_mandible=WomMaData[,,39]

# Check that ply files and raw coordinates align
plotspec(mean_specimen_mesh_cranium,mean_coords_cranium, centered = F, asp=T) 
plotspec(mean_specimen_mesh_mandible,mean_coords_mandible, centered = F, asp=T)

#warp reference mesh to mean shape - warped ply files are already in the Data/Processed folder
#wombat_cranium_meanwarp- warpRefMesh(mean_specimen_mesh_cranium, mean_coords_cranium, mshape_cranium, centered=F) # Warps the ref mesh to the mean shape
#wombat_mandible_meanwarp <- warpRefMesh(mean_specimen_mesh_mandible, mean_coords_mandible, mshape_mandible, centered=F) # Warps the ref mesh to the mean shape

#save mean meshes
#open3d(); shade3d(wombat_cranium_meanwarp); writePLY("../Data/Processed/wombat_cranium_meanwarp.ply",withColors=T, format=("ascii"))#Export the warped mean specimen as a ply file; if you want to import this back into R, it needs to be formatted to ascii ply in meshlab
#open3d(); shade3d(wombat_mandible_meanwarp); writePLY("../Data/Processed/wombat_mandible_meanwarp.ply",withColors=T, format=("ascii"))

#or open meshes 
wombat_cranium_meanwarp<-read.ply("../Data/Processed/wombat_cranium_meanwarp.ply")
wombat_mandible_meanwarp<-read.ply("../Data/Processed/wombat_mandible_meanwarp.ply")

#for plotting coloured landmarks on surface mesh

lm_types_cranium=read.csv("../Data/Raw/landmarktypes_cranium_wombat.csv")
lm_types_mandible=read.csv("../Data/Raw/landmarktypes_mandible_wombat.csv")

#separate colour vectors are needed because the code assigns the first landmark type it encounters withe the first colour in the vector. 

col_cranium <- c("blue", "orange", "green")
col_mandible <- c("orange","blue", "green")

#plotting coloured landmarks on surface mesh for partitions; change to mandible as needed

PartLevels= unique(land_data$cranium$landmarkgroups[,2])
Colours <- c("blue", "orange", "green")

Part=list()
###Subsetting the partition levels
for(i in 1:length(PartLevels)){
  Part[[i]]<-which (land_data$cranium$landmarkgroups[,2] == PartLevels[[i]])
} 

###plotting the spheres- if you get only white, increase the radius.       
open3d()
for (i in 1:length(PartLevels)){
  spheres3d(mshape_cranium[Part[[i]],1], mshape_cranium[Part[[i]],2], mshape_cranium[Part[[i]],3], col=Colours[i], lit=TRUE,radius = 0.001, asp=F)
  
}

shade3d(wombat_cranium_meanwarp, color="grey")



#change from here between cranium and cranium as needed

PartLevels= unique(lm_types_mandible$x)


Part=list()

###Subsetting the partition levels
for(i in 1:length(PartLevels)){
  Part[[i]]<-which (lm_types_mandible[,3] == PartLevels[[i]])
} 

###plotting the spheres- if you get only white, increase the radius       
open3d()
for (i in 1:length(PartLevels)){
  spheres3d(mshape_mandible[Part[[i]],1], mshape_mandible[Part[[i]],2], mshape_mandible[Part[[i]],3], col=col_mandible[i], lit=TRUE,radius = 0.001, asp=F)
  
}

shade3d(wombat_mandible_meanwarp, color="grey")

```

#Supplementary Figure 3 - multivariate allometry

```{r}
cranium_gdf=geomorph.data.frame(coords=land_data$cranium$procrustes$coords, Csize=land_data$cranium$procrustes$Csize)
AllomCranium=procD.lm(coords~Csize, data=cranium_gdf, method="RegScore", logsz=FALSE)


mandible_gdf=geomorph.data.frame(coords=land_data$mandible$procrustes$coords, Csize=land_data$mandible$procrustes$Csize)
AllomMandible=procD.lm(coords~Csize, data=mandible_gdf, method="RegScore", logsz=FALSE)


## Allometry plots

gg.color.hue <- function(n) {
    hues = seq(15, 375, length = n + 1)
    hcl(h = hues, l = 65, c = 100)[1:n]
  }

#pdf("../Manuscript/Figures/Allom_Suppfigure.pdf", height=6, width=12)
par(mfrow = c(1,2)) 
plotAllometry(AllomCranium, size=cranium_gdf$Csize, method="RegScore", bg = gg.color.hue (length(levels(cranium$Species))) [cranium$Species], pch=21, cex=1.5, logsz = FALSE);

legend ("topleft", "Cranium", bty="n", cex=1.2);

legend("bottomright", c("Common", "Southern HN", "Northern HN"), pt.bg=(gg.color.hue(3)),pch=21, bty="n", cex=1.2 )

plotAllometry(AllomMandible, size=mandible_gdf$Csize, method="RegScore", bg = gg.color.hue (length(levels(mandible$Species))) [mandible$Species], pch=21, cex=1.5, logsz = FALSE);

legend ("topleft", "Mandible", bty="n", cex=1.2);

legend("bottomright", c("Common", "Southern HN", "Northern HN"), pt.bg=(gg.color.hue(3)),pch=21, bty="n", cex=1.2 )

#dev.off()


```