---
title: "Main analyses"
author: "Vera Weisbecker"
bibliography: references.bib
date: "`r Sys.Date()`"
output:
  html_document:
    fig_width: 12
    fig_height: 6
---

```{r, message = FALSE, warning = FALSE}
## Loading the libraries (and installing if necessary)
if(!require(knitr)) install.packages("knitr")
if(!require(geomorph)) install.packages("geomorph")
source("../Functions/utilities.R")
set.seed(42)

## Loading and renaming the land_data objects
load("../Data/Processed/wombat.Rda")
AllWombats <- land_data
load("../Data/Processed/wombat_lasiorhinus.Rda")
HN<-land_data
load("../Data/Processed/wombat_krefftii.Rda")
NHNW <- land_data
load("../Data/Processed/wombat_latifrons.Rda")
SHNW <- land_data
load("../Data/Processed/wombat_ursinus.Rda")
CW <- land_data

## Read the classifiers
cranium <- read.csv("../Data/Raw/classifier_cranium_wombat.csv")
mandible <- read.csv("../Data/Raw/classifier_mandible_wombat.csv")

## Remove the NHNW_B34 (juvenile)
nhnw_b34 <- which(mandible$ID == "NHNW_B34")
mandible <- mandible[-nhnw_b34, ]
```

# Preparing the data

## Partition classifiers according to species

<!-- Full_classifier_Table_ is the tables of mandibular or cranial classifiers (e.g. read.csv("../Data/Raw/classifier_cranium_wombat.csv")); spp is a vector of species (e.g. c("CW", "SHNW", "NHNW")) -->

```{r, message = FALSE, warning = FALSE}
## Creating the vectors holding the species names
species <- c("CW", "SHNW", "NHNW")
sp_holders <-c("Common wombat", "SHN wombat", "NHN wombat")

## Assigning the right class to each species (for the cranium)
for (i in 1:length(unique(cranium$Species))) {
    ## Extracting the class for each species   
    class <- cranium[cranium$Species == sp_holders[i],]
    ## Preparing the sp name for giving names to outputs
    sp_name <- paste(species[i], deparse(substitute(cranium)), "class", sep = "_")
    ## Assigning the class to the names
    assign(sp_name, class)
}

## Assigning the right class to each species (for the mandible)
for (i in 1:length(unique(mandible$Species))) {  
    ## Extracting the class for each species   
    class <- mandible[mandible$Species == sp_holders[i],]
    ## Preparing the sp name for giving names to outputs
    sp_name <- paste(species[i], deparse(substitute(mandible)), "class", sep = "_")
    ## Assigning the class to the names
    assign(sp_name, class)
}
```

## Checking for shape outliers

The following session evaluates the shape of the outliers

```{r, message = FALSE, warning = FALSE}
## set up the gridpars for plotRefToTarget to be passed to rgl
points <- gridPar(pt.size = 0.7)
```

This following section detects outliers (the specimen beyond the upper and lower quartiles) and then visually checks whether these individuals have any obvious wrongly placed landmarks.
When multiple individuals are outliers, each are checked in turn (replacing `ID` in `X$Y$procrustes$coords[, , ID]` by the specimens' ID; where `X` and `Y` are respectively the species object and the species dataset).

#### The Common wombat

The cranium:

```{r, message = FALSE, warning = FALSE, eval = FALSE}
## Plot the procrustes distances from mean for specimens to identify the ones beyond
## the upper/lower quartiles
plotOutliers(CW$cranium$procrustes$coords) 
## CW_M11 (individual 14,11 are out)

## Use plotRefToTarget with lollipop graphs to check if outlier has landmark error;
## Replace number according to which individuals have been identified as outliers
plotRefToTarget(CW$cranium$procrustes$consensus, CW$cranium$procrustes$coords[, , 14],
                method = "vector", gridPars = points)
## No evidence of landmarking error
```

The mandible:

```{r, message = FALSE, warning = FALSE, eval = FALSE}
## Plot the procrustes distances from mean for specimens to identify the ones beyond
## the upper/lower quartiles
plotOutliers(CW$mandible$procrustes$coords)
## individuals 20 and 11 are out

## Use plotRefToTarget with lollipop graphs to check if outlier has landmark error;
## Replace number according to which individuals have been identified as outliers
plotRefToTarget(CW$mandible$procrustes$consensus, CW$mandible$procrustes$coords[, , 20],
                method = "vector", gridPars = points)
## No evidence of landmarking error
```

#### The Northern Hairy Nose Wombat

The cranium:

```{r, message = FALSE, warning = FALSE, eval = FALSE}
## Plotting the outliers
plotOutliers(NHNW$cranium$procrustes$coords)
## NHNW_M11 (individual 21 is out)

## Plotting the outliers
plotRefToTarget(NHNW$cranium$procrustes$consensus, NHNW$cranium$procrustes$coords[, , 21],
                method = "vector", gridPars = points)
## No evidence of landmarking error
```

The mandible:

```{r, message = FALSE, warning = FALSE, eval = FALSE}
## Plotting the outliers
plotOutliers(NHNW$mandible$procrustes$coords)
## Specimen 3 is out

## Plotting the outliers
plotRefToTarget(NHNW$mandible$procrustes$consensus, NHNW$mandible$procrustes$coords[, , 3],
                method = "vector", gridPars = points)
## No evidence of landmarking error
```

#### The Southern Hairy Nose Wombat

The cranium:

```{r, message = FALSE, warning = FALSE, eval = FALSE}
## Plotting the outliers
plotOutliers(SHNW$cranium$procrustes$coords)
## 21, 20 and 5 are out)

## Plotting the outliers
plotRefToTarget(SHNW$cranium$procrustes$consensus, NHNW$cranium$procrustes$coords[, , 5],
                method = "vector", gridPars = points)
## No evidence of landmarking error
```

The mandible:

```{r, message = FALSE, warning = FALSE, eval = FALSE}
## Plotting the outliers
plotOutliers(SHNW$mandible$procrustes$coords)
## No outliers!
```


# Assessing range of Centroid sizes

Here we look at the effect of species on the centroid sizes (i.e. are the centroid sizes different within each species?):

```{r, message = FALSE, warning = FALSE, fig.height = 6, fig.width = 6}
## Placeholder list
centroid_size <- list()

## separating the centroid sizes
for (i in 1:length(levels(cranium$Species))) {
    centroid_size[[i]] <- list()
    centroid_size[[i]] <- AllWombats$cranium$procrustes$Csize[
                              cranium$Species == levels(cranium$Species)[i] ]
}

## Naming the centroid sizes
names(centroid_size) = levels(cranium$Species)

## Centroid size ranges
max(centroid_size$`Common wombat`) - min(centroid_size$`Common wombat`)
max(centroid_size$`NHN wombat`) - min(centroid_size$`NHN wombat`)
max(centroid_size$`SHN wombat`) - min(centroid_size$`SHN wombat`)

## Getting all the centroid sizes
c_sizes <- c(centroid_size$`Common wombat`, centroid_size$`NHN wombat`, centroid_size$`SHN wombat`)
## Assigning each one to a species factor
species_c_size <- as.factor(c(cranium$Species))

## Plotting the results
boxplot(c_sizes ~ species_c_size, names = levels(cranium$Species), range = 1, ylab = "Centroid Size")

## Running the anova
c_size_aov <- aov(c_sizes ~ species_c_size)
c_size_tukey <- TukeyHSD(c_size_aov, "species_c_size")
```

We can summarise these results as follows:

```{r}
## Renaming the factors for the tukey results
tukey_results <- c_size_tukey$species_c_size
rownames(tukey_results) <- gsub("1", names(centroid_size)[1], rownames(tukey_results))
rownames(tukey_results) <- gsub("2", names(centroid_size)[2], rownames(tukey_results))
rownames(tukey_results) <- gsub("3", names(centroid_size)[3], rownames(tukey_results))

## The anova results
kable(summary(c_size_aov)[[1]], digits = 3, caption = "anova results from centroid size ~ species")

## The posthoc differences
kable(tukey_results, digits = 3,
      caption = "Posthoc differences from the Tukey HSD test (centroid size ~ species)")
```

There is a significant difference in centroid size between the southern hairy nose wombat's centroid size and the other wombats (common and northern hairy nose).

# Allometry Analyses

We can run the allometry tests using the `handle.procD.formula` wrapper function:

```{r, message = FALSE, warning = FALSE, result = 'hide'}
## List of all the data
all_data <- list("NHNW" = NHNW, "SHNW" = SHNW, "CW" = CW)

## Placeholder for the results
allom_output <- list()

## Running the allometry tests
for (i in 1:length(all_data)){
    for (k in 1:length(all_data[[1]])){
        ## Result placeholder
        allom_output[i][[k]] <- list()
        
        ## Procrustes allometry test (wrapper around the procD.allometry function)
        allom_output[[i]][[k]] <- handle.procD.formula(coords ~ Csize, all_data[[i]][[k]]$procrustes,
                                                       procD.fun = procD.allometry, iter = 1000,
                                                       print.progress = FALSE)
    }
}

## Naming the results
names(allom_output) <- names(all_data)

## Naming the partitions
for(i in 1:length(all_data)) {
    names(allom_output[[i]]) <- names(all_data[[i]])
}
```

And we can display them as follows:

```{r, echo = FALSE}
kable(allom_output[[1]][[1]]$aov.table, digits = 3,
    caption = "Procrustes ANOVA for size (allometry) for the northern hairy nose wombat's cranium.")
kable(allom_output[[1]][[2]]$aov.table, digits = 3,
   caption = "Procrustes ANOVA for size (allometry) for the northern hairy nose wombat's mandible.")
kable(allom_output[[2]][[1]]$aov.table, digits = 3,
    caption = "Procrustes ANOVA for size (allometry) for the southern hairy nose wombat's cranium.")
kable(allom_output[[2]][[2]]$aov.table, digits = 3,
   caption = "Procrustes ANOVA for size (allometry) for the southern hairy nose wombat's mandible.")
kable(allom_output[[3]][[1]]$aov.table, digits = 3,
              caption = "Procrustes ANOVA for size (allometry) for the common wombat's cranium.")
kable(allom_output[[3]][[2]]$aov.table, digits = 3,
              caption = "Procrustes ANOVA for size (allometry) for the common wombat's mandible.")
```

## Allometry analysis without the patches

We can also run the allometry analysis on the data without the patches


```{r, message = FALSE, warning = FALSE, result = 'hide'}
## Placeholder for the results
allom_output_no_patch <- list()

## Running the allometry tests
for (i in 1:length(all_data)){
    for (k in 1:length(all_data[[1]])){
        ## Result placeholder
        allom_output_no_patch[i][[k]] <- list()
        
        ## Procrustes allometry test (wrapper around the procD.allometry function)
        allom_output_no_patch[[i]][[k]] <- handle.procD.formula(coords ~ Csize,
                                                       all_data[[i]][[k]]$procrustes_no_patches,
                                                       procD.fun = procD.allometry, iter = 1000,
                                                       print.progress = FALSE)
    }
}

## Naming the results
names(allom_output_no_patch) <- names(all_data)

## Naming the partitions
for(i in 1:length(all_data)) {
    names(allom_output_no_patch[[i]]) <- names(all_data[[i]])
}
```

And we can display them as follows:

```{r, echo = FALSE}
kable(allom_output_no_patch[[1]][[1]]$aov.table, digits = 3,
    caption = "Procrustes ANOVA for size (allometry) for the northern hairy nose wombat's cranium (no patches).")
kable(allom_output_no_patch[[1]][[2]]$aov.table, digits = 3,
   caption = "Procrustes ANOVA for size (allometry) for the northern hairy nose wombat's mandible (no patches).")
kable(allom_output_no_patch[[2]][[1]]$aov.table, digits = 3,
    caption = "Procrustes ANOVA for size (allometry) for the southern hairy nose wombat's cranium (no patches).")
kable(allom_output_no_patch[[2]][[2]]$aov.table, digits = 3,
   caption = "Procrustes ANOVA for size (allometry) for the southern hairy nose wombat's mandible (no patches).")
kable(allom_output_no_patch[[3]][[1]]$aov.table, digits = 3,
              caption = "Procrustes ANOVA for size (allometry) for the common wombat's cranium (no patches).")
kable(allom_output_no_patch[[3]][[2]]$aov.table, digits = 3,
              caption = "Procrustes ANOVA for size (allometry) for the common wombat's mandible (no patches).")
```


# Other procD analyses

## Make the geolocation and sex subsets

The following analysis don't contain the northern hairy nose wombats because there is no variation in location and no sex assignment in the dataset.

```{r, message = FALSE, warning = FALSE}
## Making the classifiers list
all_data <- list("SHNW" = SHNW, "CW" = CW)
all_classifiers <- list("SHNW" = list("cranium" = SHNW_cranium_class,
                                      "mandible" = SHNW_mandible_class),
                          "CW" = list("cranium" = CW_cranium_class,
                                      "mandible" = CW_mandible_class)
                        )
```

Creating the geolocation subsets and the centroid size per geolocation subsets:

```{r}
## Results placeholder
coords_geo_output <- c_size_geo_output <- coords_sex_output <- c_size_sex_output <- list()

## Getting the coordinates for specimens with Geolocation
for (i in 1:length(all_data)){
    ## Result placeholder
    list() ->
        coords_geo_output[[i]] ->
        c_size_geo_output[[i]] ->
        coords_sex_output[[i]] ->
        c_size_sex_output[[i]]

    ## Getting the coordinates for each classifier
    for (k in 1:length(all_classifiers[[i]])){
        ## Geolocation coordinates
        coords_geo_output[[i]][[k]] <- all_data[[i]][[k]]$procrustes$coords[, ,
                                       as.character(all_classifiers[[i]][[k]]$Geolocation) == "Yes"]
        ## Geolocation centroid size
        c_size_geo_output[[i]][[k]] <- all_data[[i]][[k]]$procrustes$Csize[
                                       as.character(all_classifiers[[i]][[k]]$Geolocation) == "Yes"]
        ## Sex coordinates
        coords_sex_output[[i]][[k]] <- all_data[[i]][[k]]$procrustes$coords[ , ,
                                       !is.na(all_classifiers[[i]][[k]]$Sex)]
        ## Sex centroid size
        c_size_sex_output[[i]][[k]] <- all_data[[i]][[k]]$procrustes$Csize[
                                       !is.na(all_classifiers[[i]][[k]]$Sex)]

    }
    ## Naming the element
    names(all_data[[i]]) ->
      names(coords_geo_output[[i]]) ->
      names(c_size_geo_output[[i]]) ->
      names(coords_sex_output[[i]]) ->
      names(c_size_sex_output[[i]])
}

names(all_data) ->
  names(coords_geo_output) ->
  names(c_size_geo_output) ->
  names(coords_sex_output) ->
  names(c_size_sex_output)  
```

Creating the coordinates subsets and the centroid size subsets for geolocation and sex:

```{r}
## Results placeholder
geo_gdfs <- list()
sex_gdfs <- list()

## Creating the geomorph dataframe for both species and datasets
for (i in 1:length(coords_geo_output)){
    ## Results placeholder
    geo_gdfs[[i]] <- list()
    sex_gdfs[[i]] <- list()
    
    for (k in 1:length(all_classifiers[[i]])){
      ## Geolocation data frame object
      geo_gdfs[[i]][[k]] <- geomorph.data.frame(
        coords = coords_geo_output[[i]][[k]],
        Csize = c_size_geo_output[[i]][[k]],
        Temp = all_classifiers[[i]][[k]][all_classifiers[[i]][[k]]$Geolocation=="Yes",]$Temperature,
        Aridity = all_classifiers[[i]][[k]][all_classifiers[[i]][[k]]$Geolocation=="Yes",]$Aridity
        )

      ## Sex data frame object
      sex_gdfs[[i]][[k]] <- geomorph.data.frame(
                          coords = coords_sex_output[[i]][[k]],
                          Csize = c_size_sex_output[[i]][[k]],
                          Sex = all_classifiers[[i]][[k]]$Sex[!is.na(all_classifiers[[i]][[k]]$Sex)]
                          )
  }
  ## Naming the elements
  names(geo_gdfs[[i]]) <- names(sex_gdfs[[i]]) <- names(all_data[[i]])
}
names(geo_gdfs) <- names(sex_gdfs) <- names(all_data) 

## Making sure the dimension names are correct
# expect_equal(match(dimnames(geo_gdfs$SHNW$cranium$coords)[[3]],names(geo_gdfs$SHNW$cranium$Csize)),
#              seq(1:length(names(geo_gdfs$SHNW$cranium$Csize))))
```


## Running procD.lm analyses

We can run the three following linear models for each species and each datasets:

 * coordinates ~ temperature
 * coordinates ~ sex + centroid size
 * centroid size ~ sex

```{r, message = FALSE, warning = FALSE}
## Results placeholder
coord_temp_lm <- coord_sex_lm <- csize_sex_lm <- list()

## Running all the linear models not for Northern hairy-nosed wombats because of no sufficient information
for (i in 1:length(geo_gdfs)){
    ## Results placeholder
    coord_temp_lm[[i]] <- coord_sex_lm[[i]] <- csize_sex_lm[[i]] <- list()
        
    for (k in 1:length(geo_gdfs[[2]])){
        ## Running the coordinates ~ temperature models
        coord_temp_lm[[i]][[k]] <- procD.lm(coords ~ Temp, print.progress = FALSE,
                                            data = geo_gdfs[[i]][[k]], iter = 1000)
        ## Running the coordinates ~ sex + centroid size models
        coord_sex_lm[[i]][[k]] <- procD.lm(coords ~ Sex + Csize, print.progress = FALSE,
                                           data = sex_gdfs[[i]][[k]], iter = 1000)
        ## Running the centroid size ~ sex models
        csize_sex_lm[[i]][[k]] <- procD.lm(Csize ~ Sex, print.progress = FALSE,
                                           data = sex_gdfs[[i]][[k]], iter = 1000)
    }
    ## Renaming the models
    names(geo_gdfs[[i]]) ->
        names(coord_temp_lm[[i]]) ->
        names(coord_sex_lm[[i]]) ->
        names(csize_sex_lm[[i]])
  }

names(coord_temp_lm) <- names(coord_sex_lm) <- names(csize_sex_lm) <- names(geo_gdfs)
```


And we can display them as follows:

```{r, echo = FALSE}
## coordinates ~ temperature
kable(coord_temp_lm[[1]][[1]]$aov.table, digits = 3,
    caption = "Procrustes linear model for coordinates as a function of temperature for the southern hairy nose wombat's cranium.")
kable(coord_temp_lm[[1]][[2]]$aov.table, digits = 3,
    caption = "Procrustes linear model for coordinates as a function of temperature for the southern hairy nose wombat's mandible.")
kable(coord_temp_lm[[2]][[1]]$aov.table, digits = 3,
    caption = "Procrustes linear model for coordinates as a function of temperature for the common wombat's cranium.")
kable(coord_temp_lm[[2]][[2]]$aov.table, digits = 3,
    caption = "Procrustes linear model for coordinates as a function of temperature for the common wombat's mandible.")

## coordinates ~ sex + centroid size
kable(coord_sex_lm[[1]][[1]]$aov.table, digits = 3,
    caption = "Procrustes linear model for coordinates as a function of sex and centroid size for the southern hairy nose wombat's cranium.")
kable(coord_sex_lm[[1]][[2]]$aov.table, digits = 3,
    caption = "Procrustes linear model for coordinates as a function of sex and centroid size for the southern hairy nose wombat's mandible.")
kable(coord_sex_lm[[2]][[1]]$aov.table, digits = 3,
    caption = "Procrustes linear model for coordinates as a function of sex and centroid size for the common wombat's cranium.")
kable(coord_sex_lm[[2]][[2]]$aov.table, digits = 3,
    caption = "Procrustes linear model for coordinates as a function of sex and centroid size for the common wombat's mandible.")

## centroid size ~ sex 
kable(coord_sex_lm[[1]][[1]]$aov.table, digits = 3,
    caption = "Procrustes linear model for centroid size as a function of sex for the southern hairy nose wombat's cranium.")
kable(coord_sex_lm[[1]][[2]]$aov.table, digits = 3,
    caption = "Procrustes linear model for centroid size as a function of sex for the southern hairy nose wombat's mandible.")
kable(coord_sex_lm[[2]][[1]]$aov.table, digits = 3,
    caption = "Procrustes linear model for centroid size as a function of sex for the common wombat's cranium.")
kable(coord_sex_lm[[2]][[2]]$aov.table, digits = 3,
    caption = "Procrustes linear model for centroid size as a function of sex for the common wombat's mandible.")
```

# Shape disparity comparison

Here we look at the disparity measured as the relative sum of the diagonal of the covariance of the ordination scores (Procrustes variance), and we measure the difference between groups using permutation tests.


```{r, message = FALSE, warning = FALSE}
## Create gdfs with species for the data frames
cranium_gdf  <- geomorph.data.frame(coords = AllWombats$cranium$procrustes$coords,
                                    Csize = AllWombats$cranium$procrustes$Csize,
                                    species = as.factor(cranium$Species))
mandible_gdf <- geomorph.data.frame(coords = AllWombats$mandible$procrustes$coords,
                                    Csize = AllWombats$mandible$procrustes$Csize,
                                    species = as.factor(mandible$Species))

## Running the analyses on the gdf objects (Disparity within groups compared to that group's mean)
cranium_disparity <- morphol.disparity(coords ~ species, groups = ~species, iter = 1000, seed = 42,
                                       data = cranium_gdf, print.progress = FALSE)

## Disparity within groups compared to that group's mean
mandible_disparity <- morphol.disparity(coords ~ species, groups = ~species, iter = 1000, seed = 42,
                                        data = mandible_gdf, print.progress = FALSE)
```

And we can display the results as follows:

```{r, echo = FALSE}
## Cranium disparity values
kable(data.frame("Procrustes variance" = cranium_disparity[[1]]), digits = 5, caption = "Cranium disparity (procrustes variance) per groups.")
## Cranium disparity differences
kable(cranium_disparity[[2]], digits = 5, caption = "Cranium pairwise absolute differences between variances between groups.")
## Cranium disparity differences p
kable(cranium_disparity[[3]], digits = 5, caption = "Cranium pairwise absolute differences between variances between groups (p-values).")

## Mandible disparity values
kable(data.frame("Procrustes variance" = mandible_disparity[[1]]), digits = 5, caption = "Mandible disparity (procrustes variance) per groups.")
## Mandible disparity differences
kable(mandible_disparity[[2]], digits = 5, caption = "Mandible pairwise absolute differences between variances between groups.")
## Mandible disparity differences p
kable(mandible_disparity[[3]], digits = 5, caption = "Mandible pairwise absolute differences between variances between groups (p-values).")
```


# 2-Block Partial Least Squares analysis (2BPLS)


## Reduce coordinate datasets to only those that have mandibles and crania

```{r, message = FALSE, warning = FALSE}
## Data list
all_data <- list("NHNW" = NHNW, "SHNW" = SHNW, "CW" = CW)
## Classifiers list
all_classifiers <- list(
                    "NHNW" = list("cranium" = NHNW_cranium_class, "mandible" = NHNW_mandible_class),
                    "SHNW" = list("cranium" = SHNW_cranium_class, "mandible" = SHNW_mandible_class),
                    "CW" = list("cranium" = CW_cranium_class, "mandible" = CW_mandible_class)
                   )

## Checking whether the cranial and mandibular coordinates datasets match
matched_coords <- reduce.check(all_data, all_classifiers)
# expect_true(all(unlist(matched_coords[[2]])))
```

## Run 2BPLS

Depending on processor speed, this can take a while, so we recommend just loading the results which are saved in the `Results` folder.
The results can be run using the following code chunk:


```{r, message = FALSE, warning = FALSE, eval = FALSE}
## 2BPLS for the southern hairy nose wombats
SHNWTBPLS <- two.b.pls(matched_coords[[1]]$SHNW$cranium ,matched_coords[[1]]$SHNW$mandible,
                       iter = 1000) 
save(SHNWTBPLS, file = "../Data/Results/SHNWTBLS.Rda")

## 2BPLS for the northern hairy nose wombats
NHNWTBPLS <- two.b.pls(matched_coords[[1]]$NHNW$cranium ,matched_coords[[1]]$NHNW$mandible,
                       iter = 1000) 
save(NHNWTBPLS, file="../Data/Results/NHNWTBLS.Rda")

## 2BPLS for the common wombats
CWTBPLS <- two.b.pls(matched_coords[[1]]$CW$cranium ,matched_coords[[1]]$CW$mandible,
                     iter = 1000) 
save(CWTBPLS, file="../Data/Results/CWTBLS.Rda")
```

Or directly loaded using this one:
 
```{r}
## 2BPLS for the southern hairy nose wombats
load("../Data/Results/SHNWTBLS.Rda")
## 2BPLS for the northern hairy nose wombats
load("../Data/Results/CWTBLS.Rda")
## 2BPLS for the common wombats
load("../Data/Results/NHNWTBLS.Rda")
```

## Subseting data to run correlations of PLS score with csize and allometry

```{r, message = FALSE, warning = FALSE}
## Getting the list of 2BPLS results
all_TBPLS <- list("NHNW" = NHNWTBPLS, "SHNW" = SHNWTBPLS, "CW" = CWTBPLS)

## Placeholder lists
pls_pca_c_size <- TBPLS_scores <- pc_scores <- centroid_size <- pls_pca <- merged_pls <- list()

## Creating the subsets
for (i in 1:length(all_data)){
 
  ## Placeholder lists
  list() -> 
      pls_pca_c_size [[i]] ->
      TBPLS_scores[[i]] ->
      pc_scores[[i]] -> 
      centroid_size[[i]] ->
      pls_pca [[i]] ->
      merged_pls[[i]]

  ## Create data frames for subsequent merging. 
  for (k in 1:length(all_data$NHNW)){

    ## TBPLS needs to contain i(species), an address for k (x/yscores as per cranium/mandible, and a
    ## designation of how long the first singular vector goes, i.e. the species number); the 5+k i
    ## the "address" of cranial/mandibular scores in the TBPLS data frames
    TBPLS_scores[[i]][[k]] <- as.data.frame(all_TBPLS[[i]][[5+k]]
                                            [c(1:length(dimnames(all_TBPLS[[i]][[5+k]])[[1]]))],
                                            row.names = dimnames(all_TBPLS[[i]][[5+k]])[[1]])
    ## Get the PC scores
    pc_scores[[i]][[k]] <- as.data.frame(all_data[[i]][[k]]$ordination$x[,1])

    ## Get the centroid sizes
    centroid_size[[i]][[k]] <- as.data.frame(all_data[[i]][[k]]$procrustes$Csize)  
    
    ## Combine the pls data
    pls_pca[[i]][[k]] <- merge(TBPLS_scores[[i]][[k]], pc_scores[[i]][[k]],
                               by = "row.names", all.x = TRUE)
    ## Renaming the rows
    row.names(pls_pca[[i]][[k]]) <- pls_pca[[i]][[k]]$Row.names

    ## Combined the merged PLS
    merged_pls[[i]][[k]] <- merge(pls_pca[[i]][[k]], centroid_size[[i]][[k]],
                                  by = "row.names", all.x = TRUE)
    ## Renaming the rows
    row.names(merged_pls[[i]][[k]]) <- merged_pls[[i]][[k]]$Row.names

    ## Remove the two first columns
    merged_pls[[i]][[k]] <- merged_pls[[i]][[k]][,-(1:2)]
    ## Renaming the columns
    colnames(merged_pls[[i]][[k]]) <- c("PLS_Scores", "PC1","Csize")
  }

  ## Add the species names
  names(merged_pls)[[i]] <- names(all_data)[[i]]
  
  ## Add the dataset name (cranium/mandible)
  names(merged_pls[[i]]) <- names(all_data[[1]])
}


## A manual visual check if the above merger is accurate; print the below and the merged_pls
SHNWTBPLS_mandible_Scores <- as.data.frame(SHNWTBPLS$YScores[c(1:19)],
                                           row.names = dimnames(SHNWTBPLS$YScores)[[1]])

SHNWPC1 <- as.data.frame(SHNW$mandible$ordination$x[,1])
SHNWCsize <- as.data.frame(SHNW$mandible$procrustes$Csize)
```

# Run the analyses

<!-- TG: this part I actually have no clue what's happening, I've left blank comments to fill someday -->

```{r, message = FALSE, warning = FALSE}
## Placeholder lists
pca_pls_lm <- c_size_pls_lm <- list()
 
##
for (i in 1:length(merged_pls)){
    ## Placeholders
    pca_pls_lm[[i]] <- c_size_pls_lm[[i]] <- list()
    
    ##
    for (k in 1: length(merged_pls[[1]])){
        pca_pls_lm[[i]][[k]] <- list()
        ##
        pca_pls_lm[[i]][[k]][[1]] <- summary(lm(
                                          merged_pls[[i]][[k]]$PLS_Scores ~ merged_pls[[i]][[k]]$PC1
                                            ))
        ##
        pca_pls_lm[[i]][[k]][[2]] <- summary(lm(
                                        merged_pls[[i]][[k]]$PLS_Scores ~ merged_pls[[i]][[k]]$Csize
                                            ))
       
    }
    ## Renaming the list elements
    names(pca_pls_lm)[[i]] <- names(merged_pls)[[i]]
    names(pca_pls_lm[[i]]) <- names(merged_pls[[1]])
    names(pca_pls_lm[[i]][[k]]) <- c("pls_pca", "PLS_Csize")
}
```

