---
title: "Other analyses"
author: "Vera Weisbecker"
bibliography: references.bib
date: "`r Sys.Date()`"
output:
  html_document:
    fig_width: 12
    fig_height: 6
---

```{r, message = FALSE, warning = FALSE}
## Loading the libraries (and installing if necessary)
if(!require(knitr)) install.packages("knitr")
if(!require(geomorph)) install.packages("geomorph")
source("../Functions/utilities.R")
set.seed(42)

#loading the below loads an object called land_data, NOT "wombat"
load("../Data/Processed/wombat_krefftii.Rda")
ClassifierCr=read.csv("../Data/Raw/classifier_cranium_wombat.csv")
```

#Allometry analyses

```{r, message = FALSE, warning = FALSE}

allom.shape(land_data$cranium$procrustes)

```
#2-Block Partial Least Squares analysis
```{r, message = FALSE, warning = FALSE}


#reduce datasets down to those that have mandibles and crania
#@params GPA_coords is a file containing coords, e.g. land_data_cranium

#Classifiers


```{r}
#I want to later make this a function for "Full_Classifier_Table" but this was to help me figure out if the code works in principle
Full_Classifier_Table_Cranial=read.csv("../Data/Raw/classifier_cranium_wombat.csv")

#I want these names to go into the output
species <- c("CW", "SHNW", "NHNW")
#these are required to pull out the subsets
sp_holders <- c("Common wombat", "SHN wombat", "NHN wombat")

    for (i in length(sp_holders)){
      
    Class=Full_Classifier_Table_Cranial[Full_Classifier_Table_Cranial$Species==sp_holders[i],]
    assign(species[i],Class)#This does what I want it to do but overwrites the Class so only the last run is saved.
    
    }



```