---
title: "Extreme specimens"
author: "Thomas Guillerme"
date: "`r Sys.Date()`"
output:
  html_document:
    fig_width: 12
    fig_height: 6
---


```{r, message = FALSE, warning = FALSE}
## Loading the libraries (and installing if necessary)
if(!require(devtools)) install.packages("devtools")
if(!require(geomorph)) install.packages("geomorph")
if(!require(landvR)) install_github("TGuillerme/landvR")
source("../Functions/utilities.R")
set.seed(42)
```

## Vombatus ursinus

```{r}
## Loading the data
load("../Data/Processed/wombat_ursinus.Rda")

## The procrustes data
procrustes_cranium <- land_data$cranium$procrustes
procrustes_mandible <- land_data$mandible$procrustes

## The range selection (based on spherical coordinates radius)
range_cranium <- variation.range(procrustes_cranium, return.ID = TRUE)
range_mandible <- variation.range(procrustes_mandible, return.ID = TRUE)

## The min/max specimens
dimnames(procrustes_cranium$coords)[[3]][range_cranium$min.max]
dimnames(procrustes_mandible$coords)[[3]][range_mandible$min.max]
```

For _Vombatus usrsinus_ the min/max specimens based on the cranium are the specimens `r dimnames(procrustes_cranium$coords)[[3]][range_cranium$min.max]`.
For the mandible they are `r dimnames(procrustes_mandible$coords)[[3]][range_mandible$min.max]`.

## Lasiorhinus latifrons

```{r}
## Loading the data
load("../Data/Processed/wombat_latifrons.Rda")

## The procrustes data
procrustes_cranium <- land_data$cranium$procrustes
procrustes_mandible <- land_data$mandible$procrustes

## The range selection (based on spherical coordinates radius)
range_cranium <- variation.range(procrustes_cranium, return.ID = TRUE)
range_mandible <- variation.range(procrustes_mandible, return.ID = TRUE)

## The min/max specimens
dimnames(procrustes_cranium$coords)[[3]][range_cranium$min.max]
dimnames(procrustes_mandible$coords)[[3]][range_mandible$min.max]
```

For _Lasiorhinus latifrons_ the min/max specimens based on the cranium are the specimens `r dimnames(procrustes_cranium$coords)[[3]][range_cranium$min.max]`.
For the mandible they are `r dimnames(procrustes_mandible$coords)[[3]][range_mandible$min.max]`.

## Lasiorhinus kreftii

```{r}
## Loading the data
load("../Data/Processed/wombat_krefftii.Rda")

## The procrustes data
procrustes_cranium <- land_data$cranium$procrustes
procrustes_mandible <- land_data$mandible$procrustes

## The range selection (based on spherical coordinates radius)
range_cranium <- variation.range(procrustes_cranium, return.ID = TRUE)
range_mandible <- variation.range(procrustes_mandible, return.ID = TRUE)

## The min/max specimens
dimnames(procrustes_cranium$coords)[[3]][range_cranium$min.max]
dimnames(procrustes_mandible$coords)[[3]][range_mandible$min.max]
```

For _Lasiorhinus kreftii_ the min/max specimens based on the cranium are the specimens `r dimnames(procrustes_cranium$coords)[[3]][range_cranium$min.max]`.
For the mandible they are `r dimnames(procrustes_mandible$coords)[[3]][range_mandible$min.max]`.