---
title: "Measuring landmark variation distances"
author: "Thomas Guillerme"
date: "`r Sys.Date()`"
output:
  html_document:
    fig_width: 12
    fig_height: 6
---

I've tried to kept the variables and comments as similar as yours.
The comments starting with `#VW:` are yours.

## Inputs

Make sure you're in the right folder where the `coordinates.difference.R`, `ClassificationWomCr.csv` and `Wombat_Cranium_Lndmrk_Data.csv` files are.

 1. First load the package:

```{r}
## Loading the libraries (and installing if necessary)
if(!require(devtools)) install.packages("devtools")
if(!require(geomorph)) install.packages("geomorph")
if(!require(zoo)) install.packages("zoo")
if(!require(dispRity)) install_github("dispRity")
## Loading the function
source("coordinates.difference.R")
```

 2. Second, load the data

```{r}
#VW: Contains information about each specimen for crania or wombats
ClassifierCr <- read.csv("ClassificationWomCr.csv", header = TRUE)
## The landmarks list
landmarks <- read.csv("MovingLMs.csv")
#VW: Read data
WomCrData <- read.csv("Wombat_Cranium_Lndmrk_Data.csv", header = TRUE, row.names = 1)
#VW: Transpose rows and columns because of the format that Viewbox exports in
WomCrData <- t(WomCrData)
#VW: Remove any empty rows
WomCrData <- WomCrData[complete.cases(WomCrData),]
#VW: Make 3D Array
WomCrData <- arrayspecs(WomCrData, k = 3, p = ncol(WomCrData)/3)
```

## Analysis

### Procrustes

```{r}
#VW: The surface and curve landmarks were already slid in viewbox, hence we do not slide them here
WomCrGPA <- gpagen(WomCrData, print.progress = FALSE)
WomCrgdf <- geomorph.data.frame(WomCrGPA, species = ClassifierCr$Species)
```

### Ordination

```{r}
WomCrPCA <- plotTangentSpace(WomCrGPA$coords, axis1 = 1, axis2 = 2, warpgrids = FALSE)
```
> By the way, small coding comment (ignore me if I sound pedantic) but using `F` instead of `FALSE` is generally a bad practice: you save four letters but at a cost of a lot of potential trouble. `F` is a variable, not a boolean (`TRUE`/`FALSE`) which as `FALSE` attributed as default. But you can easily do `F <- TRUE` and `F` will now be `TRUE`! However, you can never do `FALSE <- TRUE`.

Note that the spread of the variance is pretty even past the third axis (roughly + 2% variance per axis).

After checking, these PC*i* min/max are the transposed mean procrustes coordinates plus the dot product of the min/max for the *i^{th}* PC axis and the transposed eigenvectors.

```{r}
## This is what ploTangentSpace does individually
get.pc.min.max <- function(axis, what, PCA, GPA) {
    if(length(axis) == 1) {
       output <- geomorph::arrayspecs(as.vector(t(GPA$consensus)) + c(what(PCA$pc.scores[,axis]),
                                      rep(0, ncol(PCA$pc.scores)-length(axis))) %*% t(PCA$rotation),
                                      dim(GPA$consensus)[1], dim(GPA$consensus)[2])
    } else {

        ## To check with Emma for this one

       output <- geomorph::arrayspecs(as.vector(t(GPA$consensus)) + c(apply(PCA$pc.scores, 2, what),
                                      rep(0, ncol(PCA$pc.scores)-length(axis))) %*% t(PCA$rotation),
                                      dim(GPA$consensus)[1], dim(GPA$consensus)[2])
    }
}
```

### Using the `coordinates.difference` function

Small digression to explain the function:
This function is pretty straight forward an simply extracts the coordinates of the axis plotted by `plotRefToTarget`.
The function intakes the arguments

 * coordinates: the coordinates for one species or more.
 * reference: the reference coordinates to draw the vectors from.
 * type: which type of coordinates, `"cartesian"` or `"shperical"` (there's also `"vector"` but it's not tested yet).
 * angle: the angle unit (`"degree"` or `"radian"`).

For example, if you want the coordinates of each vectors in your first ordination axis, you can go as follow:

```{r}
first_axis_max_min <- coordinates.difference(WomCrPCA$pc.shapes$PC1min, WomCrPCA$pc.shapes$PC1max,
                                             type = "cartesian")
head(first_axis_max_min[[1]])
```

So these are the simple x,y,z coordinates of the vectors.
It's possible to visualise them (here just in 2D for the example):

```{r, fig.width = 6, fig.height = 6}
## Selecting only the two first dimension
d2_matrix <- first_axis_max_min[[1]][,c(1,2,4,5)]
## And lets plot the hairs (in 2D)
par(bty = "n")
plot(NULL, xlim = range(d2_matrix[,-c(1,2)]), ylim = range(d2_matrix[,-c(3,4)]))
apply(d2_matrix, 1, function(x) lines(x = x[1:2], y = x[3:4]))
```

Not really meaningfull! But at least we've extracted them properly.

# Measuring the differences between landmarks

## The "easy" ways

Now we can calculate the length (radius) and the angles (polar and azimuth) of all this on the first PC axis:

```{r}
first_axis_max_min <- coordinates.difference(WomCrPCA$pc.shapes$PC1min, WomCrPCA$pc.shapes$PC1max,
                                             type = "spherical", angle = "degree")
head(first_axis_max_min[[1]])
```

And draw it as distributions:

```{r, fig.width = 6, fig.height = 12}
par(mfrow = c(3,1), bty = "n")
plot(density(first_axis_max_min[[1]][,1]), main = "Distribution of landmark changes",
     xlab = "radius")
plot(density(first_axis_max_min[[1]][,2]), main = "", xlab = "azimuth")
plot(density(first_axis_max_min[[1]][,3]), main = "", xlab = "polar")
```

Interestingly there's mainly small changes in length and big changes in angle.

### Actually testing the differences

If your hypothesis is "Within each species, the landmarks selected landmarks change more than the other landmarks", you can test it by using an anova, t-test or any variance comparison test (parametric or not).
For example, let say that my selected landmarks are the ones from 1 to 100 (it's probably not the case), you can go as follow:

```{r, fig.width = 6, fig.height = 6}
## Selecting the landmarks
selected_land <- which(as.numeric(landmarks[,2]) == 1)

## Selecting the changes (radii)
selected_landmark_change <- first_axis_max_min[[1]][selected_land,1]
non_sele_landmark_change <- first_axis_max_min[[1]][-selected_land,1]

## Comparing both
par(bty = "n")
boxplot(selected_landmark_change, non_sele_landmark_change, ylab = "Radius",
        xlab = "Selected/others")

## Wilcoxon test (distributions are not normal)
(wilcox_test_PC1 <- wilcox.test(selected_landmark_change, non_sele_landmark_change))
```

Because of the huge parameter, we can check if this change is actually meaningful or simply because they are many points.
We can first measure the probability of the two distributions being the same:

```{r}
## Bhattacharrya Coefficient (probability of overlap)
(bhat_PC1 <- bhatt.coeff(selected_landmark_change, non_sele_landmark_change))
```

And then simulate both this probability and the p-values for random selected landmarks

```{r, fig.width = 6, fig.height = 12}
## Randomly sample
rand.sample <- function(max_min, selected_lan) {
    ## Randomly sampling
    random_select <- sample(1:length(max_min), length(selected_lan))

    ## Wilcox test
    wilcox <- wilcox.test(max_min[random_select], max_min[-random_select])

    ## Bhattaacharrya
    bhatt <- bhatt.coeff(max_min[random_select], max_min[-random_select])

    ## Results
    return(c("BC" = bhatt, wilcox$statistic, wilcox$p.value))
}

## Doing it on 100 bootstraps
test <- replicate(100, rand.sample(first_axis_max_min[[1]][,1], selected_land))

## Function for plotting the results
plot.one.test <- function(test, obs, which, xlab = "") {
    densi <- hist(test[which,], plot = FALSE)
    plot(NULL, main = "", xlab = xlab, xlim = range(c(densi$breaks, obs)), ylim = range(c(0, densi$counts)), ylab = "Density")
    hist(test[which,], add = TRUE)
    abline(v = obs, lwd = 2, lty = 2)
}

## Plotting the results
par(mfrow = c(3,1), bty = "n")
plot.one.test(test, bhat_PC1, which = 1, xlab = "Bhattacharrya")
plot.one.test(test, wilcox_test_PC1$statistic, which = 2, xlab = "Wilcox W")
plot.one.test(test, wilcox_test_PC1$p.value, which = 3, xlab = "p value")
```

In this case, yes, the landmark on the selected change more than the other landmarks change more (in term of distance).

#### Problem
However, I would argue that this is probably wrong in a shape sense (not mentioning that the test is not appropriate here) since what we are actually measuring, is not the change in landmark position but the change in landmark variance/co-variance along the first axis of the PCA. The hypothesis tested above was: "Within each species, the first variance/covariance axis of the selected landmark change more than the other landmarks".

We need to either measure it on all the variance/covariance (all the PC axis) or simply the physical landmark position change.

## The "hard" way

### Measuring change in position with *all* the PC axis

In the same way as shown before, we can also extract the max and min on all dimensions all together:

```{r}
all_max <- get.pc.min.max(axis = 1:70, what = max, PCA = WomCrPCA, GPA = WomCrGPA)
all_max <- matrix(all_max, dim(all_max)[1], dim(all_max)[2])
all_min <- get.pc.min.max(axis = 1:70, what = min, PCA = WomCrPCA, GPA = WomCrGPA)
all_min <- matrix(all_min, dim(all_min)[1], dim(all_min)[2])
```

And then the same way we can have the polar coordinates (but in all dimensions!):

```{r}
all_axis_max_min <- coordinates.difference(all_min, all_max,
                                           type = "spherical", angle = "degree")
head(all_axis_max_min[[1]])
```

And draw it as distributions:

```{r, fig.width = 6, fig.height = 12}
par(mfrow = c(3,1), bty = "n")
plot(density(all_axis_max_min[[1]][,1]), main = "Distribution of landmark changes", xlab = "radius")
plot(density(all_axis_max_min[[1]][,2]), main = "", xlab = "azimuth")
plot(density(all_axis_max_min[[1]][,3]), main = "", xlab = "polar")
```

These are now the variation of all species along the maximum and minimum variation in all dimensions (not just one).
We can thus test the same hypothesis as before but by making it more specific: "Within each species, the variance/covariance of the selected landmarks change more than the other landmarks."

#### Testing the differences on all axis

```{r, fig.width = 6, fig.height = 6}
## Selecting the changes (radii)
selected_landmark_change <- all_axis_max_min[[1]][selected_land,1]
non_sele_landmark_change <- all_axis_max_min[[1]][-selected_land,1]

## Comparing both
par(bty = "n")
boxplot(selected_landmark_change, non_sele_landmark_change, ylab = "Radius",
        xlab = "Selected/others")

## Wilcoxon test (distributions are not normal)
(wilcox_test_all_PC <- wilcox.test(selected_landmark_change, non_sele_landmark_change))

## Bhattacharrya Coefficient (probability of overlap)
(bhat_all_PC1 <- bhatt.coeff(selected_landmark_change, non_sele_landmark_change))
```

And then simulate both this probability and the p-values for random selected landmarks

```{r, fig.width = 6, fig.height = 12}
## Doing it on 100 bootstraps
test <- replicate(100, rand.sample(all_axis_max_min[[1]][,1], selected_land))

## Plotting the results
par(mfrow = c(3,1), bty = "n")
plot.one.test(test, bhat_all_PC1, which = 1, xlab = "Bhattacharrya")
plot.one.test(test, wilcox_test_all_PC$statistic, which = 2, xlab = "Wilcox W")
plot.one.test(test, wilcox_test_all_PC$p.value, which = 3, xlab = "p value")
```

### Measuring change in position with *pre* ordination

Always in the same way as shown before, we can extract the max and min of variation in landmark change in the actual 3D word.
This has the advantage of directly measuring what we want to test the hypothesis "Within each species, the selected landmarks change more than the other landmarks" without using the variance/covariance of the landmarks as a proxy.

For that, we'll first need to detect which species are the "maximum" and "minimum" variation.
Because they are a 826x3 variables per specimen, it's not really feasible/meaningful to simply calculate a distance matrix.
Luckily (!) we are just interested in the change in length and... we have a function that can calculate that!

Now we need to find the maximum and minimum of these landmarks distributions and here's the trick:
 
 1. we know that the radius measured above are the distances from the consensus (average shape);
 2. we can measure the integral (or the area under curves if we don't want to bother defining the functions) that will represent one value for each specimen: the bigger the value, the more difference from the consensus the specimen is (the other way around, the more average).
 3. we can find which specimen is the most different from the average and then re-measure the coordinates differences from this one.
 4. and then re-measure the areas: the max value there will be the specimen the most different from the less average specimen (the max/min specimen).

```{r}
## The area function
area <- function(coordinate_diff, what = "radius") {
    y <- sort(coordinate_diff[, what])
    x <- 1:length(y)
    return(sum(diff(x) * zoo::rollmean(y,2)))
}
```
 1. The distances from consensus:

```{r}
## Distances from the consensus
diff_from_consensus <- coordinates.difference(WomCrGPA$coords, WomCrGPA$consensus,
                                             type = "spherical", angle = "degree")
```

 2. The maximum area under the curve:

```{r}
## Getting all the areas
areas_of_radius_mean <- unlist(lapply(diff_from_consensus, area))

## Finding the specimen the most different from the consensus
max_specimen <- which(areas_of_radius_mean == max(areas_of_radius_mean))
```

 3. The distances from the most different specimen:

```{r}
## Distances from the maximum
diff_from_max <- coordinates.difference(WomCrGPA$coords, WomCrGPA$coords[,,max_specimen],
                                        type = "spherical", angle = "degree")
```

 4. The maximum area under the curve (the most different from the most different from average specimen)

```{r}
## Getting all the areas
areas_of_radius_max <- unlist(lapply(diff_from_max, area))

## Finding the specimen the most different from the consensus
min_specimen <- which(areas_of_radius_max == max(areas_of_radius_max))
```

So we now have our max and min specimens (though impossible to say which one is the "bigger" or "smaller" one - but we don't care about that).

Anyway, we can now measure this distance:


```{r}
## Distances between the max/min
diff_from_max_min <- coordinates.difference(WomCrGPA$coords[,,min_specimen],
                                            WomCrGPA$coords[,,max_specimen],
                                            type = "spherical", angle = "degree")
```


```{r, fig.width = 6, fig.height = 12}
par(mfrow = c(3,1), bty = "n")
plot(density(diff_from_max_min[[1]][,1]), main = "Distribution of landmark changes",
    xlab = "radius")
plot(density(diff_from_max_min[[1]][,2]), main = "", xlab = "azimuth")
plot(density(diff_from_max_min[[1]][,3]), main = "", xlab = "polar")
```

And now we can apply the same comparisons as we did for the differences in PC scores.
The advantages however is that we are now actually measuring this differences in physical units: i.e. if the differences in zygomatic arches landmarks are now measured in units of Procrustes superimposition (which are by definition 3D Euclidean distances) rather than in ordination of variance/co-variance of landmarks position.



#### Testing the differences in the Euclidean space

```{r, fig.width = 6, fig.height = 6}
## Selecting the changes (radii)
selected_landmark_change <- diff_from_max_min[[1]][selected_land,1]
non_sele_landmark_change <- diff_from_max_min[[1]][-selected_land,1]

## Comparing both
par(bty = "n")
boxplot(selected_landmark_change, non_sele_landmark_change, ylab = "Radius",
        xlab = "Selected/others")

## Wilcoxon test (distributions are not normal)
(wilcox_test_Euclidean <- wilcox.test(selected_landmark_change, non_sele_landmark_change))

## Bhattacharrya Coefficient (probability of overlap)
(bhat_Euclidean <- bhatt.coeff(selected_landmark_change, non_sele_landmark_change))
```

And then simulate both this probability and the p-values for random selected landmarks

```{r, fig.width = 6, fig.height = 12}
## Doing it on 100 bootstraps
test <- replicate(100, rand.sample(diff_from_max_min[[1]][,1], selected_land))

## Plotting the results
par(mfrow = c(3,1), bty = "n")
plot.one.test(test, bhat_Euclidean, which = 1, xlab = "Bhattacharrya")
plot.one.test(test, wilcox_test_Euclidean$statistic, which = 2, xlab = "Wilcox W")
plot.one.test(test, wilcox_test_Euclidean$p.value, which = 3, xlab = "p value")
```


### Visual differences:

Just some plots for showing the differences (in 3D so I won't run it in this markdown file):

```{r, eval = FALSE}
gridPar = gridPar(pt.bg = "white", pt.size = 0.5)
open3d()
plotRefToTarget(WomCrPCA$pc.shapes$PC1min, WomCrPCA$pc.shapes$PC1max, method = "vector",
                gridPars = gridPar, label = FALSE, main = "Max/Min on the first PC axis")
```

```{r, eval = FALSE}
plotRefToTarget(all_min, all_max, method = "vector",
                gridPars = gridPar, label = FALSE, main = "Max/Min on all the PC axes")
```

```{r, eval = FALSE}
plotRefToTarget(WomCrGPA$coords[,,max_specimen], WomCrGPA$coords[,,min_specimen], method = "vector",
                gridPars = gridPar, label = FALSE, main = "Max/Min in \"real space\"")
```

Note how the results using only the first PC axis actually distorts the reality from the one using all the axis or the one without ordination.
I think this is a classical problem in Geometric Morphometric analysis with ordination (or any dimensionality reduction analysis for that sake): the PC axis deform reality (that's what they're supposed to do) and it is then really tempting (and often easy) to propose post-hoc explanation of what is happening along each axis.
Not saying it's a wrong method but at least they is room from discussion on which method is actually more adapted to test the actual hypothesis.

I'll be curious to see the results of the properly tested hypothesis with the three methods!


